An antisense oligonucleotide-based strategy to ameliorate cognitive dysfunction in the 22q11.2 Deletion Syndrome

Version 1

Abstract

Abstract

Adults and children with the 22q11.2 Deletion Syndrome demonstrate cognitive, social and emotional impairments and high risk for schizophrenia. Work in mouse model of the 22q11.2 deletion provided compelling evidence for abnormal expression and processing of microRNAs. A major transcriptional effect of the microRNA dysregulation is up-regulation of Emc10, a component of the ER membrane complex, which promotes membrane insertion of a subset of polytopic and tail-anchored membrane proteins. We previously uncovered a key contribution of EMC10 in mediating the behavioral phenotypes observed in 22q11.2 deletion mouse models. Here we show that expression and processing of miRNAs is abnormal and EMC10 expression is elevated in neurons derived from 22q11.2 deletion carriers. Reduction of EMC10 levels restores defects in neurite outgrowth and calcium signaling in patient neurons. Furthermore, antisense oligonucleotide administration and normalization of Emc10 in the adult mouse brain not only alleviates cognitive deficits in social and spatial memory but sustains these improvements for over two months post injection, indicating its therapeutic potential. Broadly, our study integrates findings from both animal models and human neurons to elucidate the translational potential of modulating EMC10 levels and downstream targets as a specific venue to ameliorate disease progression in 22q11.2 Deletion Syndrome.

Introduction

Adults and children with the 22q11.2 Deletion Syndrome (22q11.2DS) demonstrate cognitive, social and emotional impairments (1–3). 22q11.2 deletions are also one of the strongest genetic risk factors for schizophrenia (SCZ) (4). Previous work in a model of the 22q11.2 deletion, carrying a hemizygous 1.3-Mb deficiency on mouse chromosome 16 [Df(16)A], which is syntenic to the 1.5Mb 22q11.2 deletion [Df(16)A+I– mice] revealed a distinct behavioral and cognitive profile (5, 6). Molecular analysis of the Df(16)A+I–strain provided compelling evidence for abnormal processing of brain enriched microRNAs (miRNAs) (5, 7). The Df(16)A+I– related miRNA dysregulation is due to (i) hemizygosity of Dgcr8, a component of the “microprocessor” complex that is essential for miRNA production (8) and (ii) hemizygosity of miRNA genes residing within the deletion, including mir185. Reduction of miR-185 levels and to a lesser degree of miRNAs residing outside the deletion {such as miR-485 (7)} result in a de-repression of Mirta22IEmc10 gene (herein after referred to as Emc10), whose expression is under the repressive control of miRNAs(7). Indeed, comprehensive RNA profiling of Df(16)A+I– mice found that postnatal elevation in the expression of Emc10 gene represents a key transcriptional effect of the 22q11.2 deletion (7). Increased brain expression of Emc10 is recapitulated in Df(16)A+/− primary neurons(9) as well as in mouse models of the more common 3Mb 22q11.2 deletion (10). Other miRNA targets are dysregulated, but their levels of change are subtler and more variable. Emc10 encodes for a component of the ER membrane complex (EMC), which promotes membrane insertion and maturation of a subset of polytopic and tail-anchored membrane proteins including neurotransmitter receptors, channels, and transporters (11–18). Emc10 is a prenatally biased gene with high expression in embryonic life that gradually subsides after birth (7), a developmental pattern of expression conserved between mice, humans and nonhuman primates (19). Emc10 LoF mutation that leads to reduction of Emc10 levels rescues key cellular, cognitive and behavioral alterations in the Df(16)A+I– mice (19).

Here we show that 22q11.2 deletion results in abnormal processing of miRNAs in human neurons and in turn drives misexpression of EMC10 as previously described in animal models (5). Human EMC10 expression is elevated in neurons derived from 22q11.2 deletion carriers and reversal of EMC10 expression leads to restoration of key morphological and functional alterations linked to 22q11.2 deletions, supporting normalization of EMC10 expression as a disease-modifying intervention. Toward this end, we also show that antisense oligonucleotide (ASO)-mediated Emc10 normalization in the adult mouse brain is effective at reversing cognitive alterations. Improvements in cognition are sustained for over two months post ASO administration. The observations that ASO-mediated Emc10 reduction in adult mouse brain rescues cognitive deficits linked to 22q11.2 deletion strongly support a key contribution of Emc10 and Emc10-dependent membrane protein trafficking in mediating the effects of 22q11.2 deletions on cognitive function and pave the way towards translating these observations into potential disease-modifying therapeutic interventions.

Results

To investigate whether miRNA dysregulation and upregulation of EMC10 is also prominent in cortical neurons from patients carrying 22q11.2 deletions (Fig. 1A), we used hiPSC lines obtained from three independent 22q11.2DS/SCZ donors with a 3Mb deletion diagnosed with SCZ and matched healthy controls (Supplementary Table 1). The first patient/control pair is derived from dizygotic twins discordant for the 22q11.2DS and SCZ [Q6 (22q11.2) and Q5 (Ctrl)] (Fig. S1A-D). The second patient/control pair is derived from siblings [Q1 (22q11.2) and Q2 (Ctrl)] while the third patient/control pair is a case and age/sex matched unrelated control pair from the NIMH Repository and Genomic Resource [QR27 (22q11.2) and QR20 (Ctrl)].

EMC10 is robustly upregulated in hiPSC-derived neurons from 22q11.2 deletion carriers.

(A) Schematic diagram depicting the human chromosome 22q11.2 region. Bright grey and red horizontal bars indicate the two most common hemizygous genomic deletions found in the 22q11.2 deletion syndrome. The location of the coding genes and non-coding RNAs (miRNAs, underlined) are shown for chromosome 22q11.2. The microprocessor DGCR8 (DiGeorge Syndrome Critical Region Gene 8) and mir185 are shown in bold. (B) Cortical marker TBR1 and pan-neuronal marker TUJ1 expression in cortical neurons as detected by immunocytochemistry at day 13 of differentiation. TBR1 (red), TUJ1 (green) and DAPI (blue) expression are shown. Scale bars: 100 μm. (C) Volcano plot showing differentially expressed miRNAs (DEmiRs) in cortical neurons at day 8 of differentiation. Significant DEmiRs (P value < 5%) are shown above red line; Q5 (Ctrl) n=3, Q6 (22q11.2) n=3. 153/133 miRNAs were significantly up– and downregulated in Q6 (22q11.2) hiPSC-derived cortical neurons, respectively. 22q11.2 deletion region residing miRNAs miR-185, miR-1286 and miR-1306 are highlighted. (D-F) Consistent upregulation of EMC10 mRNA in Q6 (22q11.2) line derived cortical neurons as assayed by qRT-PCR at (D) day 8 (P= 0.031; Q5: n=3, Q6: n=3), (E) day 20 (P= 0.0478; Q5: n=4, Q6: n=4) and (F) day 34 (P= 0.0358; Q5: n=3, Q6: n=3) of differentiation. (G) Western blot analysis showing upregulated EMC10 protein levels in Q6 (22q11.2) line derived cortical neurons at day 8 of differentiation. Tubulin was probed as a loading control. (H) Immunofluorescence images of NGN2 generated cells. Representative images of NGN2-iNs at DIV21 from Q5 (Ctrl) and Q6 (22q11.2) hiPSC lines identified via EGFP fluorescence and immunostained for neuronal dendrite marker MAP2 and the nuclear marker DAPI. Scale bars=100 µm. (I-K) qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21. (I) Upregulation of EMC10 mRNA in Q6 (22q11.2) line derived neurons compared to the healthy control line Q5 (P=0.0222). Q5 (Ctrl) n=4, Q6 (22q11.2) n=4. (J) Upregulation of EMC10 mRNA in Q1 (22q11.2) patient line compared to healthy control line Q2 (P=0.0441). Q2 (Ctrl) n=5 and Q1 (22q11.2) n=7. (K) Upregulation of EMC10 mRNA in QR27 (22q11.2) patient line compared to healthy control line QR20 (P=0.0414). QR20 (Ctrl) n=5 and QR27 (22q11.2) n=5. Data are presented as mean ±SEM, unpaired two-tailed t-test, *P<0.05, **P<0.01.

We examined whether 22q11.2 deletion results in abnormal processing of miRNAs in human neurons as we have previously described in animal models (5). We performed parallel small RNA/miRNA sequencing on DIV8 differentiated human cortical neurons from the sibling (Q5/Q6) pair derived using an approach that combines small-molecule inhibitors to repress SMAD and WNT signaling pathways to promote CNS fate (20). We confirmed the efficiency of differentiation using immunohistochemistry (IHC) and gene expression assays, which indicated the anticipated increase of TUJ1/TBR1 positive derived neurons and downregulation of embryonic stem cell marker OCT4 (Fig. 1B and Fig. S1E-G). We identified a number of mature miRNAs dysregulated in response to the 22q11.2 deletion (Fig. 1C, Fig. S2A, B, Supplementary Table S2). As a validation of our approach, we observed the expected downregulation of miRNAs miR-185, miR-1286 and miR-1306 that reside in the 22q11.2 locus (three other predicted 22q11.2 miRNAs, miR-649, miR-3618 and miR-4761 were not detected in DIV8 neurons) (Fig. S2A). Among miRNAs located outside the 22q11.2 region, we note downregulation of miRNAs such as miR-137 as well as miR-134 and several other members form the largest placental mammal-specific miRNA gene cluster miR379-410 (Fig. S2B) that have been previously implicated in neuronal development, differentiation and function (5, 21–29). We used the miRNA-target interaction network tool miRNet 2.0 (30) to perform target enrichment and network analysis for the dysregulated miRNAs and performed GO term enrichment analysis on this target interaction network. Affected biological processes were prominently centered on cell division and intracellular protein transport (Fig. S3A) whereas cellular components were associated with the nucleus and the perinuclear region (endoplasmic reticulum and Golgi apparatus) of the cytoplasm (Fig. S3B).

In addition to 22q11.2 deletion region miRNAs, lower abundance of miRNAs in cases is likely due to haploinsufficiency of the DGCR8 gene and is expected to result in upregulation of target genes. To identify candidate miRNA target genes we performed an unbiased evaluation of the transcriptional responses using bulk RNA sequencing on RNA collected from DIV8 differentiated cortical neurons derived from the patient (Q6) and the corresponding healthy dizygotic twin (Q5) line (Fig. S4A). We observed the expected downregulation of genes within the 22q11.2 locus in patient neurons (Supplementary Table 3). Further, RNA and protein expression characterization confirmed the reductions in the abundance of the 22q11.2 locus residing genes DGCR8 and RANBP1 (Fig. S4B, C). Among the differentially expressed genes (DEGs) 2094 were downregulated and 1937 were upregulated. As expected EMC10 expression was elevated in patient neurons while expression of other EMC subunits (EMC1-4, EMC6-9) detected in our DIV8 sequencing data did not show significant differences. GO term enrichment analysis on downregulated DEGs identified significantly altered biological processes centered on neurogenesis, neuronal development, and differentiation (Fig. S4D). Among the upregulated DEGs, the GO terms enriched were related to neuronal development as well as neuronal cilia assembly and structure (Fig. S4E).

Intersection of predicted targets of downregulated miRNAs and up-regulated DEGs identified 774 predicted targets of downregulated miRNAs (Fig. S4F, Supplementary Table 4) including EMC10. Notably, functional annotation revealed that predicted targets of downregulated miRNAs include genes that modulate neuronal development and are associated with GO terms such as endoplasmic reticulum and endomembrane system of neurons (Fig. S4G, H).

qRT-PCR assays confirmed a robust and significant upregulation of EMC10 levels in RNA extracted from cortical neurons derived from hiPSCs of the Q5/Q6 pair through SMAD/WNT signaling inhibition, at three distinct stages of in vitro maturation (Fig. 1D-F). Additionally, this upregulation was confirmed in protein extracts from cortical neurons at day 8 of differentiation (Fig. 1G). To examine whether transcriptional EMC10 upregulation is independent of the neuronal derivation method, we generated neurons via inducible expression of Neurogenin-2 (NGN2), a widely used protocol that generates a robust population of excitatory neurons (NGN2-iNs) within 3 weeks (31–34). MAP2 staining was used to demonstrate the successful neuronal differentiation of the hiPSC lines (Fig. 1H). qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21 confirmed transcriptional EMC10 upregulation in three independent pairs of patient and sex-/age matched healthy control lines (Fig. 1I-K). Taken together our results highlight a reproducible and robust upregulation of EMC10 in neurons derived from patients with 22q11.2 deletions, which is independent of the derivation method. It is noteworthy that in addition to monolayer cultures, EMC10 shows significant upregulation along the excitatory neuron lineage (radial glia, intermediate progenitors and excitatory neurons) but not in astrocytes, choroid or interneuron lineage cells, in patient forebrain organoids generated by the same hiPSCs lines used in the present study (35).

We have previously shown that upregulation of the murine orthologue of Emc10 is primarily due to downregulation of miR-185 and to a lesser degree of miR-485 (7). Both conserved and non-conserved binding sites at the 3’UTR of human EMC10 are predicted in silico for both miRNAs (Fig. 5SA). Consistently, the observed upregulation in the levels of EMC10 gene is accompanied by a robust reciprocal decrease in the levels of the miRNA precursor of miR-185 at DIV8 as indicated both by our miRNA sequencing analysis (Fig. 2A, Supplementary Table 2) and follow up qRT-PCR assays (Fig. 2A). miRNA precursor of miR-485 showed a more modest non-significant reduction in its abundance (Fig. 2B). Notably, overexpression of miR-185 and miR-485 using miRNA mimics in human cortical neurons at DIV10 resulted in reduction of EMC10 expression levels in both the healthy control (Q5, Fig. 2C) and patient line (Q6, Fig. 2D). Furthermore, inhibition of endogenous miR-185 and miR-485 in the control line by using specific miRNA inhibitors increased EMC10 expression level (Fig. 2E) confirming the predicted conserved functionality of miR-185 and miR-485 miRNA binding sites in EMC10. It is worth noting that in addition to miR-185, non-conserved binding sites at the 3’UTR of human EMC10 are predicted in silica for two additional downregulated miRNAs residing within the 22q11.2 locus, miR-1286 and miR-1306 (Supplementary Table S). The functionality of these miRNA binding sites in EMC10 and whether they contribute to the observed elevation of its expression in human neurons remains to be determined. Taken together our results confirm that miRNA dysregulation emerges in human neurons as a result of the 22q11.2 deletion and in turn drives misexpression of genes primarily involved in intracellular membrane and protein trafficking-related processes required for neuronal development and maturation. Among them, EMC10 represents a major downstream effector of the 22q11.2-linked miRNA dysregulation.

Altered miRNA expression in hiPSC-derived cortical neurons from 22q11.2 deletion carriers.

(A) Precursor miRNA expression level of miR-185 (P= 0.0038), predicted to target EMC10, are downregulated in Q6 (22q11.2) cortical neurons as assayed by qRT-PCR (Q5: n=3, Q6: n=3). (B) Precursor miRNA expression level of miR-485 (P= 0.0622), predicted to target EMC10, are downregulated in Q6 (22q11.2) cortical neurons as assayed by qRT-PCR (Q5: n=3, Q6: n=3). (C-E) miR-185 and miR-485 modulate EMC10 in human iPSC-derived cortical neurons. (C) qRT-PCR quantification shows reduced expression levels of EMC10 mRNA in Q5 (Ctrl) line derived cortical neurons at day 10 of differentiation transfected with miR-185 [one-way ANOVA, F (2, 6) = 6.079, P=0.0361; post hoc Bonferroni, P=0.0366] or miR-485 [post hoc Bonferroni, P=0.0464] mimics at day 8 of differentiation. Expression levels in miR-185 or miR-485 mimic-treated neurons were normalized to expression levels under scramble mimic controls treatment (n= 3, each treatment). (D) qRT-PCR quantification shows reduced expression levels of EMC10 mRNA in Q6 line-derived cortical neurons transfected with miR-185 [one-way ANOVA, F (3, 13) = 7.167, P=0.0044; post hoc Tukey, P=0.0345] or miR-485 [post hoc Tukey, P=0.0251] or a combination of both miRNA mimics [post hoc Tukey, P=0.0020]. Expression levels in miR-185, miR-485 or the combination of both mimic-treated neurons were normalized to expression levels under scramble mimic controls treatment. Ctrl mimic n=5, miR-185 mimic n=5, miR-485 mimic n=4 and miR185+miR485 mimics n=3. (E) qRT-PCR quantification shows increased expression levels of EMC10 mRNA in Q5 line derived cortical neurons transfected with miRNA inhibitors miR-185 [one-way ANOVA, F (2, 6) = 11.94, P=0.0081; post hoc Bonferroni, P=0.0057]or miR-485 [post hoc Bonferroni, P=0.0491] at day 8 of differentiation. Expression levels in miR-185 or miR-485 inhibitor-treated neurons were normalized to expression levels under scramble miRNA inhibitor controls treatment (n=3, each treatment). Data are presented as mean ±SEM, unpaired two-tailed t-test or one-way ANOVA as indicated, *P<0.05, **P<0.01.

To investigate the relevance of EMC10 de-repression in the development and function of patient neurons, we generated derivatives of the Q6 patient hiPSC line carrying either heterozygous (Q6/EMC10HET) or homozygous (Q6/EMC10HOM) EMC10 LoF mutations using standard CRISPR/Cas9 editing approaches (Fig. S6A). Mutations were confirmed by sequencing (Fig. S6A, lower panel) and karyotyping confirmed normal chromosome complement (Fig. S6B). We confirmed reduced expression levels of 22q11.2 gene RANBP1 by western blot in both derivative hiPSC lines (Fig. S6C) whereas stem-cell marker NANOG and OCT4 were equally expressed in all lines assayed by qRT-PCR (Fig. S6D, E). EMC10 mRNA and protein levels were reduced by ∼50% in the Q6/EMC10HET hiPSC line and abolished in the Q6/EMC10HOM line (Fig. S6F, G). It is noteworthy that we did not observe an upregulation of EMC10 mRNA levels in the Q6 hiPSC lines (Fig. S6F), a finding likely attributed to the general low expression level of miR-185 and miR-485 in hiPSCs (36). Indeed both miRNAs are developmentally regulated and show increased expression levels during neuronal development (https://ethz-ins.org/igNeuronsTimeCourse/) (37). Additional characterization of hiPSC-derived NGN2-iNs (Fig. 3A), conclusively demonstrated a reduction (Q6/EMC10HET) or elimination (Q6/EMC10HOM) of EMC10 mRNA (Fig. 3B). Expression assays of a panel of cell type-specific markers did not reveal significant differences between NGN2-iNs from the Q6 patient line and both derivative lines, indicating that gene editing has no adverse effect on neuronal differentiation (Fig. S6H).

Reduction of EMC10 levels restores defects in neurite outgrowth and calcium signaling in neurons from 22q11.2 deletion carriers.

(A) Representative images of NGN2-iNs at DIV21 from Q6/EMC10HET and Q6/EMC10HOM hiPSC lines identified via EGFP fluorescence and immunostained for neuronal dendrite marker MAP2 and the nuclear marker DAPI. Scale bars=50 µm. (B) qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21. EMC10 expression is normalized to near WT levels level in Q6/EMC10HET line (P=0.166) and abolished in Q6/EMC10HOM line (P<0.0001). Q5 (Ctrl) n=4, Q6 (22q11.2/SCZ) n=4, Q6/EMC10HET n=5 and Q6/EMC10HOM n=7. (C-G) Neuronal morphology analysis in Q5, Q6 (22q11.2), Q6/EMC10HET and Q6/EMC10HOM neurons. (C) Representative images of traced neurons. (D) Total neuronal length is reduced in Q6 line (P=0.0044) and restored in Q6/ EMC10HET (P=0.0253) and Q6/EMC10HOM line (P=0.0001). (E) Reduction in number of branch points/cell in Q6 (P=0.0195) is restored in the Q6/EMC10HET (P=0.0134) and Q6/EMC10HOM lines (P<0.0001). (F) Reduction in the total number of dendrites/cells in Q6 (P=0.0202) is reversed in the Q6/EMC10HET (P=0.0166) and Q6/EMC10HOM lines (P=0.0005) (G) The number of primary dendrites per cell is unchanged. Q5 (Ctrl) n=14, Q6 (22q11.2) n=16, Q6/EMC10HET n=19 and Q6/EMC10HOM n=16 neuronal cells. (H-I) Defects in cytoplasmic calcium signaling in Q6 (22q11.2) neurons are reversed in Q6/EMC10HET and Q6/EMC10HOM lines. (H) Changes in Fluo4-AM fluorescence signal intensity in response to 75mM KCl in Q5 (Ctrl), Q6 (22q11.2). Q6/EMC10HET and Q6/EMC10HOM hiPSC-derived neurons at DIV38. Q5 vs. Q6 (KS D=0.5405, P<0.0001), Q6 vs. Q6 EMC10HET (KS D=0.5556, P<0.0001) and Q6 vs. Q6 EMC10HOM (KS D=0.5676, P<0.0001). (I) Quantification of KCl-induced Fluo4 intensity peak amplitude (ΔF) demonstrates a reduction in Q6 line (P<0.0001) that is reversed in Q6/EMC10HET (P<0.0001) and Q6/EMC10HOM lines (P<0.0001). Q5 (Ctrl) n=70, Q6 (22q11.2/SCZ) n=37, Q6/EMC10HET n=82, Q6/EMC10HOM n=97 neuronal cells. (J) Heatmap (left) showing the expression of differentially regulated 237 genes in Q5 (Ctrl) and Q6 (22q11.2) that are normalized in the Q6/EMC10HET NGN2-iNs at DIV21 (n=3 per genotype). Gene Ontology (GO) biological process (BP) terms (right) associated with the up– and down-regulated genes normalized in the Q6/EMC10HET NGN2-iNs. (K) Heatmap (left) showing the expression of differentially regulated 382 genes in Q5 (Ctrl) and Q6 (22q11.2) that are normalized in the Q6/EMC10HOM NGN2-iNs at DIV21 (n=3 per genotype). Gene Ontology (GO) biological process (BP) terms (right) associated with the up– and down-regulated genes normalized in the Q6/EMC10HOM NGN2-iNs. Data are presented as mean ±SEM, unpaired two-tailed t-test or Kolmogorov–Smirnov test as indicated, *P<0.05, **P<0.01, ***P<0.001, ****P<0.0001.

Df(16)A+/− mice show impaired formation of dendrites in deep layer cortical neurons, which are faithfully recapitulated in primary neuronal cultures and are partially reversed by reduction of Emc10 levels (7). We asked whether impaired dendritic formation is also observed in human neuronal cultures from patients with 22q11.2 deletions and whether reduction of EMC10 levels could prevent such morphological alterations during neuronal maturation. We employed monolayer neuronal cultures of NGN2-iNs. Neuronal cells were fixed at DIV21 of differentiation, immuno-stained, traced and key indices of dendritic architecture were quantified (see Materials and Methods). Our analysis confirmed a reduced dendritic complexity in mutant neurons as reflected in total neuronal length, the number of branch points and the total number of dendrites per cell (Fig. 3C-F). The number of primary dendrites per cell were unchanged (Fig. 3G) in accordance with previous findings from the murine 22q11.2 deletion model where only subtle changes were detected in the number of primary neurons (7). Importantly, we found that reduction or elimination of EMC10 expressian restored to WT levels neuronal length and branchpoints.

Our previous evaluation of Ca2+ homeostasis perturbations caused by 22q11.2 deletions using Ca2+ imaging on primary neurons from Df(16)A+I–mice revealed a significantly lower amplitude of Ca2+ elevation following KCl evoked depolarization (9). This impairment was replicated in human cortical neurons from patients with 22q11.2 deletions (38) and shown to be partially restored by exogenous expression of DGCR8, indicating a potential role of miRNA dysregulation. Using the green-fluorescent calcium indicator Fluo-4 and time-lapse microscopy we confirmed a decrease in the amplitude of Ca2+ rise following KCl evoked depolarization, in patient (Q6) derived NGN2-iNs at DIV37/38 compared to the healthy twin (Q5) (Fig. S7A, B). We asked whether reduction of EMC10 levels could reverse such alterations. Notably, the observed defect in Ca2+ signaling were reversed in both Q6/EMC10HET and Q6/EMC10HOM NGN2-iNs as demonstrated by the increased amplitude of Ca2+ rise following depolarization (Fig. 3H, I). The observation that reduction of EMC10 levels fully restores the Ca2+ signaling deficits observed in patient neurans suggests that miRNA-dependent elevation in EMC10 levels may interfere with one or more sources of intracellular Ca2+ and a wide range of calcium-dependent processes.

Differentially expressed genes are often organized into functional groups or pathways based on their known biological roles. We used transcriptional profiling as an indirect measure of cellular pathways affected by the reduction of EMC10 levels by identifying genes differentially expressed between the parental Q6/Q5 lines whose expression differences are abolished or nearly abolished in either Q6/EMC10HET or Q6/EMC10HOM NGN2-iNs (Fig. 3J, K). We identified 237 (111 downregulated and 126 upregulated) and 382 (193 downregulated and 189 upregulated) such DEGs, respectively (Supplementary Table 6). In both cases, functional annotation analysis indicated highest enrichment scores for terms related to nervous system development as well as an enrichment in GO terms relevant to neuronal generation and differentiation. Intersection of “rescued” genes in Q6/EMC10HET and Q6/EMC10HOM NGN2-iNs identified 103 shared DEGs (Supplementary Table 7). Protein-protein interaction (PPI) analysis highlighted a functional cluster including 30 of these genes, such as the SCZ-associated genes PCDHA2 (39), RBFOX1 (40) and RGS4 (41, 42) (Fig. 5SA), involved in nervous system development (Fig. 5SB). It should be noted that the beneficial effect of elimination of EMC10 expression is consistent with previous findings indicating that lack of EMC10 does not compromise EMC assembly (43) and implying an auxiliary or modulatory role of EMC10 in the EMC function.

Taken together, our analysis of neurons from 22q11.2 deletion carriers indicate that elevation of EMC10 expression disrupts their development and maturation in a way similar to observations in murine neurons, and support normalization of EMC10 expression as a disease-modifying intervention. While our previous work has shown that constitutive genetic reduction of Emc10 levels rescues key cognitive and behavioral alterations in the Df(16)A+I–mice, translating these observations into therapeutic interventions requires demonstration that it is the sustained elevation of EMC10 throughout the adult life that interferes with the underlying neural processes rather than an irreversible impact on brain maturation during early development. Toward this end, we first investigated whether restoration of Emc10 levels in the brain of adult (2– 4-month-old) Df(16)A+I– mice is effective at reversing cognitive alterations (5, 19). Specifically, we examined the effects of Emc10 reduction in adult brain on SM, a cognitive domain robustly and reproducibly affected in adult Df(16)A+I– mice (6, 19, 44). Notably, SM deficits are also present in juvenile Df(16)A+I– mice as early as postnatal day 22 (Fig. S9A, B), underscoring the severity of this phenotype, which emerges during early adulthood. Deficits in social cognition are present in individuals with 22q11.2 deletions and use of rodent tasks that evaluate SM can serve as a useful proxy of the human condition. Impaired SM in Df(16)A+I– mice is fully restored by constitutive genetic reduction of Emc10 levels (19). To manipulate the expression of the Emc10 gene in adult Df(16)A+I– mice we used a Emc10 conditional ‘knockout-first’ design by conducting a Flp– and Cre-dependent genetic switch strategy (Fig. S10A). Parental Emc10+I- tm1a mice were crossed to a germline Flp mouse line that activates global Flp function and leads to the deletion of the frt-flanked sequence(s) in the offspring. The Emc10tm1c offspring from this cross carry a laxP flanked WT Emc10 allele and are essentially WT. To achieve temporal control of Emc10 expression we used an inducible UBC-Cre/ERT2 mouse line that activates global Cre function upon tamoxifen (TAM) treatment. This approach enables postnatal normalization of Emc10 expression at its endogenous locus preserving Emc10 expression within its physiological levels. We used UBC-Cre/ERT2 mice in crosses to generate compound Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I–mice. These mice have two WT Emc10 copies upregulated, as expected in the Df(16)A background, until TAM-induced Cre expression deletes the tagged Emc10 allele. We used oral gavage to deliver TAM and implement Cre-mediated Emc10 deletion during adulthood (postnatal day 56-70). Corn oil treatment served as a control. Behavioral analysis was performed on the following four groups: Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I+ mice treated with TAM (WT + TAM), Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I– mice treated with TAM (Df(16)A+I–+ TAM), Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I+ mice treated with corn oil vehicle (WT + oil), and Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I– mice treated with corn oil vehicle (Df(16)A+I– + oil). Investigation of the efficiency of Cre-mediated deletion in brain lysate preparations from prefrontal cortex (PFC) (Fig. S10B, C), hippocampus (HPC) (Fig. S10D, E) and cerebellum (CB) (Fig. S10F) confirmed that upon TAM treatment, Emc10 mRNA and protein levels were restored to near WT levels in the adult brain of Df(16)A+I– mice. As expected, Df(16)A+I– + oil mice showed impaired SM performance compared to WT + oil control littermates, which was fully rescued upon TAM treatment. Specifically, upon reintroduction of a familiar juvenile mouse Df(16)A+I–+ TAM mice showed a strong reduction in social interaction, indicative of intact SM, comparable to TAM-treated WT littermates and significantly different from Df(16)A+I– mice treated with corn oil (Fig. S10G). The intact SM of the Df(16)A+I– + TAM mice was further evident in analysis of difference score (Fig. S10H) compared to the corn oil-treated Df(16)A+I–mice. By contrast, Df(16)A+I–mice hyperactivity in the open field was not affected upon TAM treatment consistent with our previous results from constitutive genetic reduction of Emc10 levels (Fig. S10I). Overall, our findings indicate that restoration of Emc10 levels in adult Df(16)A+I–mice has substantial effects on ameliorating cognitive deficits, indicating a broad window of therapeutic opportunity and highlighting Emc10 as a promising target for postnatal therapeutic interventions.

We explored the translation potential of this finding by employing transient injection of single stranded ASOs targeting the mouse gene as dictated by their demonstrated efficacy as a therapeutic modality in preclinical models (45–50) and clinical studies of neurodevelopmental disorders (NDDs) or neurodegenerative disorders (51–53). Over 300 chimeric 2′-O-methoxyethyl/DNA gapmer ASOs were generated and screened for Emc10 mRNA reduction in 4T1 cells via electroporation (Fig. S11A). Lead ASOs were then confirmed in a dose-response assay (Fig. S11B), including the lead ASO (1081815, herein referred to as Emc10ASO1), which targets intron 2 of mouse Emc10 (Fig. 4A). Emc10ASO1 was selected for subsequent studies, as it was effective in lowering Emc10 expression both in vitro and in vivo. Specifically, following transient intracerebral ventricular (ICV) injection in the posterior ventricle of 8 weeks old WT mice, Emc10ASO1 effectively suppressed the levels of Emc10 mRNA (Fig. S11C) and protein (Fig. S11D) in both left and right HPC compared to WT mice treated with a control ASO (CtrlASO1) without complementarity in the mouse. Analysis of Gfap and Aif1 expression did not reveal any changes (Fig. S11E, F) suggesting lack of astroglial and microglial activation upon Emc10ASO1 injection. Emc10ASO1 injected mice showed normal gait and no signs of behavioral toxicity. IHC analysis using an antibody that selectively recognizes the phosphorothioate backbone verified a robust diffusion primarily in HPC and to a lesser degree in surrounding brain areas. Colocalization with the neuronal marker NeuN and glial fibrillary acidic protein (GFAP) confirmed accumulation in hippocampal neurons as well as GFAP-expressing astrocytes (Fig. 4B). Analysis of Df(16)A+I–mice treated by intraventricular injection at 8 weeks of age showed that Emc10ASO1 effectively lowered hippocampal Emc10 mRNA to nearly WT levels 3 weeks post-injection resulting in normalization of Emc10 expression (Fig. 4C, left panel). By contrast, consistent with the pattern of ASO distribution, we did not observe a significant reduction of Emc10 expression levels in the prefrontal cortex (PFC) of Df(16) +I– mice treated with Emc10ASO1 (Fig. 4C, right panel). In addition to targeted assays, we performed bulk RNA-sequencing analysis of CtrlASO1 and Emc10ASO1-treated Df(16)A+I–mice and WT littermates to evaluate the effect of Emc10ASO1 treatment on the hippocampal transcriptome profile. In the CtrlASO1 treated group (Fig. 4D, left panel), we observe the tripartite differential gene expression signature characteristic of Df(16)A+I– mice: upregulation of Emc10 and non-coding RNAs (pri-forms of miRNAs and long non-coding RNAs (Fig. 4D, left panel and inset, Log2Fold Change = 0.5) as well as the expected downregulation of genes included within the Df(16)A deficiency. In the Emc10ASO1 treated group (Fig. 4D, right panel and inset), Emc10 is no longer upregulated in Df(16)A+I–mice while non-coding RNAs remain upregulated, and genes included in the deficiency are robustly downregulated. Apart from Emc10, seven other genes (Mir9-3hg, Plxnd1, Cd68, Mir22hg, Gm28439, Adgre1, and Tnn) are significantly upregulated in Df(16)A+I– mice in the CtrlASO1 but not in the Emc10ASO1-treated group (Supplementary Table SS). We used the Bowtie mapping tool (54) to align short sequencing reads on both genomic and transcript sequence to assess whether these expression changes represent potential off-target effects of the Emc10ASO1 in the mouse transcriptome. Emc10ASO1 exclusively aligned with full complementarity to an intronic region in the Emc10 gene (Fig. 4A) providing additional support for high target specificity. The observed changes might represent downstream effects of Emc10 level reduction or reflect expression variability due to low expression levels of the upregulated genes.

ASO-mediated suppression of murine Emc10 in vivo.

(A) Mouse Emc10 gene map plot (ENSMUST00000118808) showing the Emc10ASO1 target site. (B) Mouse brains collected three weeks post ICV injection were stained with an ASO antibody (green), counterstained with neuronal marker NeuN (red) and nuclear stain Hoechst (blue). A robust and uniform ASO diffusion (top panel) is observed in the HPC. No signal is detected in saline injected mice (bottom panel). Overlap with NeuN (yellow, top-right panel) confirms presence in neuronal cells. Accumulation in glial cells, specifically GFAP labeled astrocytes is also observed (middle-right panel and inset; ASO in green, GFAP in red, and Hoechst in blue). Images are taken with 4x, 20x and 40x objectives. (C) qRT-PCR analysis shows Emc10ASO meditated normalization of Emc10 mRNA levels in the HPC of Df(16)A+I- mice (left panel). Significant upregulation of Emc10 mRNA expression levels is seen in CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (2, 29) = 11.65, P<0.001; post hoc Tukey, P= 0.001]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+I–compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P= <0.001). CtrlASO-treated WT mice: n=9 (5 males, 4 females), CtrlASO-treated Df(16)A+I mice: n=11 (5 males, 6 females), and Emc10ASO-treated Df(16)A+I mice: n=12 (7 males, 5 females). qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+I- (right panel) compared to WT mice [one-way ANOVA, F (2, 16) = 4.253, P=0.0330; post hoc Tukey, P= 0.0385]. ASO injection does not normalize Emc10 expression levels in PFC of Df(16)A+I mice injected with Emc10ASO (post hoc Tukey vs CtrlASO treated Df(16)A+I mice, P= 0.7642). CtrlASO treated-WT male mice: n=6, CtrlASO treated-Df(16)A+I- male mice: n=5, and Emc10ASO treated-Df(16)A+I male mice: n=8. (D) Volcano plots showing up-regulation of Emc10 expression in the CtrlASO treated-Df(16)A+I- compared to the CtrlASO-treated WT mice (left panel and inset) but not in the Emc10ASO treated-Df(16)A+I- compared to the Emc10ASO-treated WT mice (right panel and inset). The expected down-regulation of genes included in the Df(16)A+I– deletion (blue) and upregulation of non-coding RNAs (ncRNAs, red) is also evident in both panels. Downregulated genes from the 22q11.2 locus (Dgcr8, Ranbp1 and Tanga2) as well as the upregulated ncRNA (miRNA-containing) gene Mirg, are highlighted. CtrlASO treated-Df(16)A+I–males: n=5, Emc10ASO treated-Df(16)A+I–males: n= 4, CtrlASO treated WT males: n=4, and Emc10ASO treated-WT males: n=4. (E-F) ASO mediated behavioral rescue of SM deficit in Df(16)A+I- mice. (E) CtrlASO-treated Df(16)A+I- mice show a robust SM deficit compared to CtrlASO-treated WT mice as indicated by the significant difference in trial 2 interaction time upon reintroduction of a familiar juvenile mouse [three-way ANOVA for Trial X Genotype X Treatment Interaction matching by trial: F (1, 38) = 9.393 P=0.0040; post hoc Tukey, P= 0.0012]. A reduction in trial 2 interaction time indicates rescue of the SM deficit in Emc10ASO-treated Df(16)A+I- compared to CtrlASO-treated Df(16)A+I- mice (post hoc Tukey, P= 0.0114). (F) A negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+I- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction: F (1, 38) =9.369, P=0.0040; post hoc Tukey, P= 0.0002]. Increase in the difference score of Df(16)A+I- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P= 0.0004). [CtrlASO-treated WT mice: n=9 (5 males, 4 females), CtrlASO-treated Df(16)A+I mice: n=11 (5 males, 6 females), Emc10ASO treated WT mice: n=9 (5 males, 4 females), and Emc10ASO-treated Df(16)A+I mice: n=13 (7 males, 6 females)]. Data are presented as mean ± SEM, *P < 0.05; **P < 0.01; ***P < 0.001.

Eight-week-old Df(16)A+I–mice and WT littermates were treated by ICV injection of Emc10ASO1 and CtrlASO1 and tested 3 weeks later in SM assays. Df(16)A+I– mice treated with CtrlASO1 showed the expected deficits in SM as reflected in the sustained high interaction time with the reintroduced familiar juvenile mouse. By contrast, Df(16)A+I– mice injected with Emc10ASO1 had significantly improved performance to levels indistinguishable from Emc10ASO1-treated WT littermates consistent with improvement of function arising from adult restoration of Emc10 levels (Fig. 4E, F). Rescue was observed in both sexes and no significant differences were seen in treatment across sexes. In control experiments, we did not observe any effects of genotype or treatment upon reintroduction of a novel juvenile mouse in trial 2 (Fig. S12A, B), strongly indicating that SM deficits are not driven by a simple task fatigue.

To evaluate the consistency and reproducibility of the ASO-mediated SM rescue, we generated additional ASOs and screened them for Emc10 mRNA reduction in viva (Fig. S13A-C). One of these ASOs (1466182, herein referred to as Emc10ASO2), which targets intron 1 of mouse Emc10 (Fig. 5A) was selected as the lead ASO for the replication analysis based on its efficacy in reducing Emc10 levels, distribution pattern, as well as the lack of any signs of astroglial/microglial activation or behavioral toxicity (see Materials and Methods). IHC analysis showed robust distribution in both HPC (Fig. S14A, top panel) and PFC (Fig. S14A, middle panel) as well as diffusion into both neuronal and non-neuronal cells (Fig. S14A, bottom panel). qRT-PCR analysis of Df(16)A+I–mice treated by intraventricular injection at 8 weeks of age showed that Emc10ASO2 effectively lowered Emc10 mRNA to nearly WT levels 3 weeks post-injection resulting in normalization of Emc10 expression in the HPC (Fig. 5B), PFC (Fig. 5C) and somatosensory cortex (Fig. 5D). To study the effects of Emc10ASO2-mediated Emc10 reduction on SM performance, 8-week-old Df(16)A+I- male mice and WT littermates were treated by ICV injection of Emc10ASO2 and CtrlASO2 and tested 3 weeks later. Df(16)A+I- mice injected with Emc10ASO2 had significantly improved SM performance to levels indistinguishable from Emc10ASO2-treated WT littermates (Fig. 5E, F).

In addition to SM deficits, Df(16)A+I–mice show a spectrum of cognitive impairments in episodic and spatial memory as reflected, for example, in impaired performance in an Y-maze-based delayed alternations task that probes short-term spatial memory (55) and contextual fear conditioning a form of associative learning test used for studying episodic learning and spatial memory (5). We proceeded to investigate the impact of ASO-mediated Emc10 reduction in the adult brain on both of these cognitive tasks. First, we confirmed that adult male Df(16)A+I–mice exhibit impaired short-term spatial memory during novelty exploration in a two-trial delayed alternation Y-maze task (Fig. 5G) as previously described for another mouse model of the 22q11.2 deletion (60). The total number of arm entries remained unchanged, indicating no alternations in locomotor activity (Fig. S14B). To determine whether reducing Emc10 expression in the brain via ASO treatment could rescue short-term spatial memory deficits, we tested a new cohort of Df(16)A+I–mice and WT littermates 3-weeks following ASO administration (Fig. 5H). ASO-treated Df(16)A+I– mice exhibited a significant improvement in delayed alternations compared to Df(16)A+I– mice treated with control ASO (Fig. 5H). Furthermore, no significant differences in total number of arm entries were observed between the groups (Fig. S14C). We confirmed the reduction of Emc10 levels in the ASO-treated animals through qRT-PCR assays of the HPC, PFC and SSC brain regions (Fig. S14D-F). In the contextual fear conditioning task, while ASO treatment was not sufficient to fully rescue the learning deficit in Df(16)A+I– mice to WT levels (Fig. 5I, right panel), there was a modest improvement in fear memory of ASO-treated Df(16)A+I– mice, since these mice did not differ significantly from the ASO-treated WT littermates. Interestingly, we have previously shown that genetic reduction of Emc10 levels in Df(16)A+I– mice resulted in only partial restoration of deficits in contextual fear memory (19). Thus our finding faithfully recapitulates results from our previous constitutive genetic rescue assays (19) and likely indicates a more limited role of Emc10 upregulation in the 22q11.2-linked fear memory deficits rather than requirement for additional treatment time or for earlier onset of Emc10 normalization.

Restoration of cognitive function in Df(16)A+I- mice using an independent Emc10 ASO.

(A) Mouse Emc10 gene map plot (ENSMUST00000118808) showing the Emc10ASO2 target site. (B) qRT-PCR analysis shows significant upregulation of Emc10 mRNA expression levels in the HPC of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (2, 19) = 20.92, P<0.0001; post hoc Tukey, P= 0.0018]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+I– compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P= <0.0001). (C) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (2, 19) = 40.75, P=<0.0001; post hoc Tukey, P=<0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in the Emc10ASO treated-Df(16)A+I–compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P=<0.0001). CtrlASO-treated WT male mice: n=6, CtrlASO-treated Df(16)A+I male mice: n=8, and Emc10ASO-treated Df(16)A+I mice: n=8. (D) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in Somatosensory Cortex (SSC) of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (2, 18) = 21.53, P=<0.0001; post hoc Tukey, P=<0.0026]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+I– compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P=<0.0001). CtrlASO-treated WT mice: n=6, CtrlASO-treated Df(16)A+I mice: n=7, and Emc10ASO-treated Df(16)A+I mice: n=8. (E) CtrlASO-treated Df(16)A+I- mice show a robust SM deficit compared to CtrlASO-treated WT mice as indicated by the significant difference in trial 2 interaction time upon reintroduction of a familiar juvenile mouse [three-way ANOVA for Trial X Genotype X Treatment Interaction matching by trial: F (1, 26) = 35.74 P<0.0001; post hoc Tukey, P= 0.0004]. A reduction in trial 2 interaction time indicates rescue of the SM deficit in Emc10ASO-treated Df(16)A+I- compared to CtrlASO-treated Df(16)A+I- mice (post hoc Tukey, P= 0.0255). [CtrlASO-treated WT male mice: n=6, CtrlASO-treated Df(16)A+I mice: n=8, Emc10ASO treated WT mice: n=8, and Emc10ASO-treated Df(16)A+I mice: n=8. (F) A negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+I- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 26) = 35.74, P<0.0001; post hoc Tukey, P<0.0001. Increase in the difference score of Df(16)A+I- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P<0.0001). (G) Y-maze task displayed memory impairments in adult Df(16)A+I- mice. Impaired short-term spatial memory in Df(16)A+I- mice shown by the reduced amount of delayed alternations (%) after a delay of 1h (P=0.0015). WT mice: n=12, Df(16)A+I mice: n=11. (H) Deficits in short-term spatial memory in the Y-maze task in CtrlASO-treated adult Df(16)A+I- mice can be rescued in Emc10ASO-treated Df(16)A+I- mice (one-way ANOVA, F(3, 25) = 6.727, P=0.0018, post hoc Tukey, P=0.0042) after 3-weeks of ASO-injection. CtrlASO-treated WT mice: n=8, Emc10ASO treated WT mice: n=8, CtrlASO-treated Df(16)A+I mice: n=8 and Emc10ASO-treated Df(16)A+I mice: n=5. (I) In a contextual fear memory assay, minimal freezing is observed on day 1 (left) with no significant changes across groups [two-way ANOVA for Genotype X Treatment interaction: F (1, 52) = 1.003, P=0.3211]. In the CtrlASO-treated group, Df(16)A+I- mice show the expected contextual fear memory deficit compared to WT mice (right panel) [one-way ANOVA, F (3, 52) = 3.524 P=0.0212; post hoc Tukey, 0.0384]. Emc10ASO treatment was not sufficient to fully rescue the learning deficit in Df(16)A+I– mice compared to CtrlASO-treated WT levels (post hoc Tukey, P= 0.1045). However, there is increased freezing in Df(16)A+I- mice injected with Emc10ASO-versus CtrlASO-treated group, which results in a non-significant difference in freezing between Emc10ASO-treated Df(16)A+I- and WT mice (two-way ANOVA for genotype x treatment F (1, 52) = 0.8524, P=0.3601; post hoc Tukey, P= 0.4676) indicating a partial rescue of the contextual fear memory deficit. CtrlASO-treated WT: n=13 (9 males, 4 females), Emc10ASO treated WT mice: n=14 (10 males, 4 females), CtrlASO-treated Df(16)A+I mice: n=15 (11 males, 4 females and Emc10ASO-treated Df(16)A+I mice: n=14 (10 males, 4 females). Unpaired students t-test, one-two-or three-way ANOVA as indicated. Data are presented as mean ± SEM, *P < 0.05, **P < 0.01, ***P < 0.01, ****P<0.0001.

The application of ASOs as a novel therapeutic strategy has seen a significant rise in recent years, in part due to their versatility in durably modifying RNA transcripts. Therefore, we investigated the longevity of ASO-mediated repression of Emc10 as an indicator of future therapeutic relevance for the treatment of 22q11.2DS. To this end, we conducted the SM and Y-maze assays on a new cohort of Df(16)A+I–mice at 3-4 weeks and 8-9 weeks post injection with Emc10ASO2 (Fig. 6A). We observed behavioral rescue in Emc10ASO2-treated Df(16)A+I– mice in the SM assay at three weeks (Fig. 6B, left panel) and in the Y-maze assay at four weeks post ASO-administration (Fig. 6C, left panel) in accordance to our previous findings (Fig. 5F, H). Remarkably, we replicated these results at 8-9 weeks post injection, demonstrating sustained behavioral rescue of SM (Fig. 6B, right panel) and spatial memory deficits (Fig. 6C, right panel). Importantly, locomotor activity remained unchanged in the Y-maze assays at both, four weeks (Fig. S14G) and nine weeks (Fig. S14H) post injection. Finally, we confirmed the downregulation of Emc10 in ASO-treated animals via qRT-PCR assays of the HPC, PFC, and SSC brain regions at ten weeks post-treatment (Fig. 6D-F). These results suggest that normalizing Emc10 expression in the brain can ameliorate social and spatial memory deficits in adult Df(16)A+I– mice in a time period of at least two months.

Sustained rescue of cognitive deficits in Df(16)A+I- mice following ASO administration.

(A) Experimental timeline of Emc10ASO2-treated adult Df(16)A+I- mice to determine long-term rescue effect of Emc10 de-repression. (B-C) Sustained ASO effect on behavioral rescue of Emc10ASO-treated Df(16)A+I- mice in HPC– and PFC-dependent tasks. (B) Rescue of SM deficit in Df(16)A+I- mice after 3– and 8-weeks of ASO injection. After 3-weeks of ASO treatment (left panel), a negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+I- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 28) = 26.20, P<0.0001; post hoc Tukey, P<0.0001]. Increase in the difference score of Df(16)A+I- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P<0.0001). After 8-weeks of ASO treatment (right panel), a negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+I- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 28) = 13.36, P=0.0010; post hoc Tukey, P=0.0012. Increase in the difference score of Df(16)A+I- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates prolonged SM rescue (post hoc Tukey, P=0.0007). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+I mice: n=7, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+I mice: n=8. (C) Y-maze task displayed rescue of memory impairments in adult Df(16)A+I- male mice 4– and 9-weeks after Emc10ASO injection. Deficits in short-term spatial memory shown by the reduced number of delayed alternations (%) in CtrlASO-treated adult Df(16)A+I- mice can be rescued in Emc10ASO-treated Df(16)A+I- mice after 4-weeks of ASO injection (left panel) [one-way ANOVA, F(3, 29) = 7.585, P=0.0007, post hoc Tukey, P=0.0193]. Deficits in short-term spatial memory in CtrlASO-treated adult Df(16)A+I- mice can be rescued in Emc10ASO-treated Df(16)A+I- mice also after 9-weeks of ASO injection (right panel) [one-way ANOVA, F(3, 29) = 6.547, P=0.0016, post hoc Tukey, P=0.0029]. CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+I mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+I mice: n=8. (D-F) Sustained Emc10 de-repression after two months of Emc10ASO injection. ASO-mediated inhibition of Emc10 in different brain regions of Df(16)A+I- mice after 10-weeks of injection. (D) qRT-PCR analysis shows significant upregulation of Emc10 mRNA expression levels in the HPC of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (3, 29) = 13.97, P<0.0001; post hoc Tukey, P=0.0331]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+I– compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P= <0.0002). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+I mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+I mice: n=8. (E) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (3, 28) = 32.47, P<0.0001; post hoc Tukey, P<0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in the Emc10ASO treated-Df(16)A+I– compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P<0.0001). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+I mice: n=7, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+I mice: n=8. (F) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in Somatosensory Cortex (SSC) of CtrlASO treated-Df(16)A+I- compared to WT mice [one-way ANOVA, F (3, 28) =23.18, P<0.0001; post hoc Tukey, P=0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+I– compared to CtrlASO treated-Df(16)A+I- mice (post hoc Tukey, P<0.0001). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+I mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+I male mice: n=7. One-or two-way ANOVA as indicated. Data are presented as mean ± SEM, *P < 0.05, **P < 0.01, ***P < 0.01, ****P<0.0001.

Panel A was created with https://BioRender.com/e28e115.

Discussion

Despite an understanding of the molecular mechanisms of 22q11.2DS, especially ones related to abnormal expression and processing of miRNAs (5, 7, 19, 38, 56), we still do not have a promising therapeutic avenue for the cognitive and neuropsychiatric symptoms associated with the 22q11.2 deletion. By leveraging our recent understanding of the molecular, cellular and behavioral consequences of 22q11.2-linked miRNA dysregulation, the present study represents an advancement towards developing a potential therapeutic strategy in two ways:

First, we show that 22q11.2 deletion results in abnormal processing of miRNAs in human neurons and in turn drives upregulation of EMC10 levels as previously described in mouse models (5). Effective reduction to near WT levels or even complete depletion of EMC10 leads to restoration of key alterations in morphological and functional neuronal maturation emerging due to 22q11.2 deletions. The miRNA regulatory mechanism underlying EMC10 upregulation and the restoration of cellular deficits are very similar to the ones we previously described in Df(16)A+I- mice, highlighting the robustness of this molecular alteration as well as the translational value of using animal models to probe the link between Emc10 upregulation and 22q11.2-linked behavioral dysfunction.

Secand, we show that normalization of Emc10 levels in adult Df(16)A+I- mouse brain, by ASO-mediated targeted knockdown, is effective in rescuing SM deficits (which emerge during postnatal development and are present as early as postnatal day 22) as well as short-term spatial memory deficits. These findings strongly suggest that at least for a subset of 22q11.2-associated cognitive deficits it is the sustained miRNA dysregulation and elevation of Emc10 throughout the adult life that interferes with the underlying neural processes rather than an irreversible impact on brain maturation during early development and demonstrate the therapeutic potential for treating a wide range of cognitive symptoms associated with 22q11.2DS.

In vivo delivery of ASOs offers a potential venue for emerging treatments of genetically driven and postnatally reversible symptoms of neurodevelopmental disorders (NDDs) focusing in reduction of culprit gene expression via sequence-specific knockdown of mRNA transcripts (57). A common challenge in efforts to employ gene-knockdown therapies for dosage-sensitive genes such as EMC10 is restricting target gene expression within optimal levels to avoid potential toxicity due to target hyper-knockdown or complete elimination (58–61). A large number of relatively rare LoF variants or potentially damaging missense variants have been identified in the human EMC10 gene among likely healthy individuals in gnomAD (60, 62) which is depleted of individuals known to be affected by severe NDDs. Taken together with our previous analysis of heterozygous Emc10 knockout mice (19), these observations strongly indicate that 50% reductions in Emc10 levels are well tolerated at the level of the organism. The highest dose for both Emc10-specific ASOs used in mutant Df(16)A+I–mice was limited to ∼300 µg and reduced Emc10 mRNA to either normal or below normal (30-50% of WT) expression while attaining full behavioral rescue. While higher dose may be required to ameliorate other behavioral deficits (58, 60), it should be noted that an acute injection of 700µg Emc10ASO in WT mice that resulted in ∼50% reduction of Emc10 levels (Fig. S13) did not cause secondary cellular and behavioral toxicity. Thus, available evidence indicates that therapeutically effective ASO-mediated normalization of EMC10 levels can be up-or down-titrated within an unequivocally safe range. Future experiments will determine whether ASO-mediated normalization of adult Emc10 levels restores additional 22q11.2 behavioral and cognitive alterations and whether earlier onset of postnatal ASO treatment can be employed to prevent the onset of behavioral deficits or to mitigate specific behavioral deficits that do not respond to adult interventions.

In addition to reduction of Emc10 expression our findings have implications for therapeutic interventions aiming to manipulate its downstream targets. In that respect, it is attractive to speculate that different EMC10 upregulation-linked phenotypes and their developmental requirements may be driven by dysregulation of distinct, individually or in combinations, downstream EMC targets. Such targets are typically multi-transmembrane domain (TMD) proteins (17, 18, 63) that contain low-hydrophobicity TMDs which are hard to insert into ER membrane and thus require the aid of EMC as a membrane insertase (11, 12, 17). Identification of neurotransmitter receptors, channels, and transporters whose biogenesis, trafficking and membrane insertion are affected by EMC10 upregulation could help establish a link between such targets and 22q11.2-related behavioral dysfunction and guide efforts to develop treatments for specific 22q11.2 deletion symptoms.

Overall, by pointing to manipulations of EMC10 expression and activity as well as of downstream targets as an attractive alternative or augmentation of currently available treatments and highlighting a broad temporal window of therapeutic and preventive opportunity for the 22q11.2 deletion-associated cognitive and behavioral symptoms, our results pave the way for developing mechanism-based therapeutic strategies, leveraging insights from both human and animal models to enhance clinical outcomes in precision medicine for neuropsychiatric disorders.

Materials and methods

Mice

We used Emc10 conditional knockout (see below) and Df(16)A+I- mice (5) in C57BL/6J background. Df(16)A+I- male mice were crossed to C57BL/6J female mice to obtain either Df(16)A+I- or WT littermates. For further details, see Supplemental Materials and Methods.

Generation of Emc10 conditional knockout compound mouse lines

To manipulate the expression of the Emc10 gene in Df(16)A+I– mice we used a Emc10 conditional ‘knockout-first’ mouse design by conducting Cre– and Flip-dependent genetic switch strategies as described earlier (64). This approach enables postnatal manipulation of Emc10 expression at its endogenous locus keeping Emc10 expression within its physiological levels. Emc10tm1a+I mice (2310044h10rik-Tm1a, MRC Harwell Institute, Oxfordshire, UK) were crossed to a germline Flp mouse line (B6.129S4-Gt(ROSA)26Sortm1(FLP1)Dym/RainJ_JAX:009086) that activates global Flp function and leads to the deletion of the frt-flanked sequence(s) in the offspring. The Emc10tm1c offspring from these cross carries a laxP flanked WT Emc10 allele. We used UBC-Cre/ERT2 mice (B6.Cg-Ndor1 Tg(UBC-cre/ERT2)1Ejb /2J_JAX:008085) in crosses to generate compound Emc10tm1c+I-; UBC-cre/ERT2; Df(16)A+I– mice.

Cell line differentiation

The cell line donors used in this study are listed in Supplementary Table S1. Human iPSC lines were maintained in mTeSR Plus medium on Matrigel coated tissue culture plate and neuronal induction of hiPSC lines into cortical neurons were conducted by using either a combination of small molecule inhibitors or NGN2 overexpression approaches. For further details, see Supplemental Material and Methods.

Genome editing of Q6(22q11.2) hiPCS line

We generated derivatives of the Q6 (22q11.2) hiPSC line carrying either heterozygous (Q6/EMC10HET) or homozygous (Q6/EMC10HOM) EMC10 LoF mutations using standard CRISPR/Cas9 genome editing approaches. For further details, see Supplemental Materials and Methods.

ASOs

Mouse Emc10-targeting ASOs used in these studies were 20 bases in length, chimeric 2’ –O-(2-methoxyethyl) (MOE)/DNA) oligonucleotides with phosphodiester and phosphorothioate linkages. The central gap of 10 deoxynucleotides is flanked on its 5’ and 3’ sides by five MOE modified nucleotides. Oligonucleotides were synthesized at Ionis Pharmaceuticals (Carlsbad, CA, USA) as described previously (65, 66). ASOs were solubilized in 0.9% sterile saline or PBS.

Stereotactic Intracerebroventricular (ICV) Injections of ASOs

ASOs were delivered to 8 weeks old mice via ICV injections using a Hamilton syringe (Hamilton Company, Reno, NV, USA) connected to a motorized Stoelting Quintessential Stereotaxic Injector QSI/53311 (Stoelting Co., Wood Dale, IL, USA). For further details, see Supplemental Materials and Methods.

Analysis of dendritic complexity

iNs were prepared on coverslips, fixed at DIV21 and immunostained for TBR1 and MAP2 to identify dendritic branches. For further details, see Supplemental Materials and Methods.

Calcium imaging

Calcium imaging was performed as previously described (67) with a few modifications. For further details, see Supplemental Materials and Methods.

Bulk RNAseq and bioinformatic analysis of mouse hippocampal samples

Total RNA was isolated from 4 WT CtrlASO1, 4 WT Emc10ASO1, 5 Df16 CtrlASO1 and 4 Df16 Emc10ASO1 treated male hippocampi. Stranded polyA+ enriched RNA sequencing libraries were prepared at the Columbia Genome Center (Columbia University, New York, USA) to generate 40 million paired-end reads on Illumina Novaseq 6000 instrument using STRYPOLYA library prep kit. For further details, see Supplemental Material and Methods.

Bulk RNAseq and small RNA/miRNAseq of hiPSC-derived cortical neurons

Total RNA was extracted from hiPSC-derived cortical neurons using the miRVana miRNA isolation kit (#AM1560, Ambion, Thermo Fisher Scientific, Waltham, MA, USA) before bulk RNAseq (paired-ended sequencing; llumina NovaSeq 6000) or miRNAseq (single-end sequencing; Illumina Hiseq 2500) were performed. For further details, see Supplemental Material and Methods.

Behavioral assays

Mice were 11-17 weeks old at the time of behavioral testing except of the social memory assay in juveniles which was performed in 3 week old animals. Behavior was assayed 3-9 weeks following surgical delivery of ASOs or 1 week following TAM/corn oil treatment. The following behavioral assays were performed in this study: Open field assay, social memory assay, contextual fear conditioning and Y-maze delayed alternation task. For further details, see Supplemental Material and Methods. The experimenter was blind to mouse genotype and treatments while performing behavioral assays and data analysis. Animals were given at least one-week intervals between behavioral tests.

Statistical Analysis

Data were analyzed using GraphPad Prism (Graphpad Software, Inc., San Diego, CA, USA). Data were evaluated as indicated, using either unpaired two-tailed t-test, Kolmogorov–Smirnov test, one-way, two-way or three-way analysis of variance (ANOVA) tests followed by post hoc Tukey’s multiple comparison test for comparisons across all groups. Data are presented as mean ±SEM. P values for each comparison are described in the figure legends.

eLife Assessment

The presented evidence is compelling given a range of complementary and mutually supportive studies. Experiments are generally robustly conducted and well-presented, supporting the claims regarding miRNA mechanisms converging on EMC10 overexpression with 22q11 Del. This is an important study that works to establish a novel antisense oligonucleotide-based approach to treating 22q11.2 deletion syndrome; the findings are likely to advance therapeutic efforts. The authors provide evidence both in vitro in patient-derived iPSCs and in vivo in a 22q11 Del mouse supporting the knockdown (KD) of EMC10 as an effective strategy for the amelioration of neuronal and behavioral deficits.

Reviewer #1 (Public review):

Summary:

This is an important and very well-presented set of experiments following up on prior work from the lab investigating knock-down (KD) of EMC10 in the restoration of neuronal and cognitive deficits in 22q11.2 Del models, including now both human iPSCs and a mouse model in vivo now with ASOs.

The valuable progress in this current manuscript is the development of ASOs, and the proof of efficacy in vivo in mice of the ASO in knock-down of EMC10 and amelioration of in vivo behavioral phenotypes.

The experiments include iPSC studies demonstrating elevations of EMC10 in a solid collection of paired iPSC lines. These studies also provide evidence of manipulation of EMC10 by overexpression and inhibition of miRNAs that exist in the 22q11 interval. The iPSC studies also nicely demonstrate the rescue of impairments with KD of EMC10 in neuronal arborization as well as KCl-induced neuronal activity. The major in vivo contributions reflect an impressive demonstration of the efficacy of two ASOs in vivo on both KD of EMC10 in vivo and through improvement in behavioral abnormalities in the 22q11 mouse in a range of different behaviors, including social behavior and learning behaviors.

Overall, there are many strengths reflected in this study, including in particular the synergy between in vitro studies in human cell models and in vivo studies in the well-characterized mouse model. The experiments are generally rigorously performed, well-powered, and nicely presented. The claims with regard to the mechanisms of EMC10 elevations and the importance of restoration of EMC10 expression to neuronal morphology and behavior are well supported by the data. The work may be further supported in future studies, by investigation of rescue by ASOs of circuit dysfunction in vivo or ex vivo through electrophysiology in the mouse model. Also, in future studies, investigation of the mechanism by which EMC10, an ER protein involved in protein processing, may function in the observed neuronal abnormalities; however, these studies are clearly for future investigations.

The potential impact of the work is found in the potential value of the ASO approach to the treatment of 22q11, or the pre-clinical evidence that knock-down of this protein may lead to some amelioration of cognitive symptoms. Overall, a very convincing and complementary set of experiments to support EMC10 KD as a therapeutic strategy.

Reviewer #2 (Public review):

Summary:

The manuscript by Thakur et. al seeks to establish a novel ASO-based approach to treat 22q11.2 deletion syndrome. Central to this thesis is that an ER membrane complex member called EMC10 is significantly increased in the disorder, which is largely attributed to the loss of miRNA-mediated repression. The authors generated three new iPSC cell lines for the disorder and showed that deletion of EMC10 rescues morphology and Ca-flux deficits. They go on to show that post-symptomatic deletion of Emc10 in mice using a conditional-off tamoxifen allele reverses social memory phenotypes. Finally, in collaboration with Ionis, they developed two new ASOs to knock down EMC10 and show that social and spatial memory phenotypes are rescued, even two months after injection.

Strengths:

In general, this represents a substantial undertaking and an impressive body of work. The experiments follow a logical progression and in most cases are well-controlled. The isolation of EMC10 effects relative to the broader miRNA disruption is viewed as impactful. The use of both genetic and ASO approaches to validate the therapeutic strategy is also viewed as highly positive. The authors' contention that EMC10 can be targeted at post-symptomatic time points to reverse 22q11.2 deletion syndrome is supported by the data. Further, they have provided a therapeutic mechanism to do so. These findings are likely to be impactful and lead to further development efforts.

Weaknesses:

The primary weaknesses of the manuscript lie in incomplete or inappropriate data analysis, as well as a failure to validate key experiments. For example, both genetic and ASO-mediated EMC10-mediated reductions are assessed at the level of mRNA, but only one experiment, in one brain region, is validated at the protein level. This brain region is the PFC, which is problematic when many of the phenotypes used have a strong hippocampal component. Likewise, the iPSC experiments make the case that excitatory neurons are central to the phenotype, but no effort is made to show that the ASOs are entering that type of neuron, or even any quantification of what percentage of cells in the target brain regions (HPC, PFC, etc.) are positive for the ASO. There is only a single image provided of staining with a phosphorothioate antibody and a claim of robust uptake, which cannot be assumed. The iPSC transcriptomics work would also benefit from a more comprehensive comparison between the EMC10 knockout lines and their parent 22q11 deletion lines. Further, there are other examples where the statistics used are either wrong (Figure 3 t-test vs ANOVA) or missing (Figure S2). These technical and analytical shortcomings make it challenging to fully interpret the data and detract from an otherwise exciting manuscript.

Author response:

The following is the authors’ response to the original reviews.

We appreciate that both reviewers found our findings significant and recognized the strength of the presented data in demonstrating the potential value of ASO-mediated Emc10 expression modulation for treating 22q11.2DS. We are grateful for the reviewers' valuable input and constructive suggestions, which we believe have significantly strengthened our manuscript. Below, we address the main points and concerns, followed by our point-by-point responses:

Evaluation of ASO-Mediated Emc10 Reduction: We appreciate the feedback and the opportunity to clarify this point. While we agree that ASO-mediated reduction of Emc10 should ideally be evaluated at both the mRNA and protein levels, we would like to emphasize that this was indeed performed in our study. Specifically, we conducted both qRT-PCR and Western Blot (WB) assays on the same animal cohort, focusing on the left and right hippocampus (rather than the PFC) following ASO injection (see Figure S11C and D). We prioritized the hippocampus for the WB assay because our primary behavioral assays and observed phenotypes in this study are strongly hippocampus-centric. This approach reflects our aim to investigate Emc10's role in the brain regions most relevant to the observed phenotypes. We hope this clarification addresses the reviewer’s concerns. While protein-level analysis would ideally complement RNA measurements, the Emc10 antibodies available were suboptimal in specificity and sensitivity, requiring substantial optimization. Additionally, challenges in obtaining sufficient high-quality protein from small regions like the hippocampus limited the use of protein detection as a standalone method. We plan to refine antibody protocols or explore alternative methods in future work. Notably, in all instances where we performed parallel protein and RNA measurements in both, mouse brain tissue and human cell lines, there was excellent concordance between the datasets, strongly suggesting that mRNA levels are a reliable indicator of Emc10 protein levels in our model.

ASO Neuronal Uptake: While ASO uptake by neurons in the brain can vary considerably depending on factors such as ASO chemistry, delivery method, target brain region, and cell type, our targeted delivery approach, ASO design optimization, and ASO screening strategy were specifically tailored to achieve uniform and efficient uptake across hippocampal and cortical regions, in both neurons and glia. The figures included in our manuscript at both low and high magnification (see Figure S14A) clearly display the extensive (over 97%) overlap of ASO-positive cells (green signal) with cells expressing the neuronal marker NeuN (red signal). While quantifying ASO-positive cells in different brain regions could add value, the robust diffusion of ASO into neurons and glia is effectively demonstrated in the current figures and indirectly supported by the robust downregulation of Emc10 in ASO-treated animals as shown by qRT-PCR assays of hippocampal and cortical brain regions.

Transcriptomic Data in Mutant EMC10 NGN2-iNs: Reduction in EMC10 levels is not expected to directly affect transcription or to broadly reorganize the differential gene expression profile of the Q6/Q5 patient/control NGN2-iN lines. Accordingly, our transcriptional profiling was not designed to assess the direct impact of EMC10 deficiency on gene expression but rather to serve as an indirect measure of cellular pathways affected by the reduction in EMC10 levels in the patient Q6 line. We aimed to identify genes and related functional pathways differentially expressed between the Q6/Q5 patient/control lines, where these expression differences are either abolished or significantly attenuated in Q6/EMC10HET or Q6/EMC10HOM NGN2-iNs.

Statistical Analysis: We have meticulously reviewed all statistical analyses in the manuscript to ensure their appropriateness and adherence to established practices. For Figure S2, we acknowledge that the statistical details were not fully specified in the figure legend, though they are provided for each miRNA in Supplemental Table S2. In the revised manuscript, we ensured that the statistical methods and corresponding values are clearly indicated for each comparison.

We are confident that the revisions outlined above, along with the point-by-point responses provided below, will significantly strengthen our manuscript and address all the concerns raised by the reviewers. We would like to express our sincere thanks to the reviewers for their valuable feedback and constructive suggestions.

Reviewer #1 (Recommendations For The Authors):

My comments here are generally limited to minor comments that reflect possible small additions or edits to the manuscript:

(1) Panel 1A is very small. Please consider making that bigger as space permits.

We have increased the panel size of Figure 1A in the revised manuscript to improve its visibility and clarity.

(2) Are you able to identify the dot that represents EMC10 in panel 1C? I understand that EMC10 is represented in Supplementary Figure 4A.

We appreciate the reviewer's observation. In Figure 1C, the volcano plot depicts differentially expressed miRNAs in the Q5/Q6 neuronal samples, as identified through miRNA-sequencing. Since EMC10, as a protein-coding gene and a downstream target of miRNA dysregulation, is not included in this analysis. However, as the reviewer correctly notes, EMC10 gene expression is represented in the volcano plot in Supplementary Figure 4A, which displays differentially expressed genes identified through bulk RNA-seq analysis of the same neuronal samples. To avoid any confusion, we have clarified the title of Figure 1C to emphasize that it represents miRNA expression changes.

(3) With regard to studies using iPSC. Some of the studies are executed across multiple distinct pairs and some are only done in a single pair. Overall, while results are coherent and often complimentary, would it be valuable for the authors to comment on experiments where studies in multiple pairs seemed particularly important, or others wherein it was less important?

We thank the reviewer for this insightful question regarding our use of multiple versus single hiPSC pairs. Our investigation began with the Q5/Q6 sibling (dizygotic twin) pair, which shares the most similar genetic background. This minimized the impact of confounding genetic factors and provided a robust foundation for testing our hypothesis that EMC10 upregulation, driven by miRNA dysregulation, is a key consequence of the 22q11.2 deletion in human neurons, thus validating our previous findings from the Df(16)A+/- mouse model (Stark et al., 2008; Xu et al., 2013). To ensure the generalizability of our findings, we incorporated additional hiPSC lines from another sibling pair as well as a case/control pair, demonstrating that EMC10 upregulation is a consistent feature of 22q11.2DS. Subsequently, we focused on the well-matched Q5/Q6 pair for detailed morphological, functional, and genetic rescue experiments. This approach allowed us to perform in-depth studies while controlling for potential genetic confounders. By using both multiple and single hiPSC pairs, we balanced the need for generalizable findings with the practical considerations of conducting technically complex and resource-intensive experiments. This strategy enabled us to provide both broad and detailed insights into the mechanisms underlying 22q11.2DS. We have modified the introductory paragraph of the Results section to better highlight this issue.

(4) While the majority of the experiments seem sufficiently powered to test the hypothesis in question in the iPSC studies, Figure 2B raises the question if the study replicates here were underpowered, and perhaps the authors might consider mentioning this, although this is a very minor comment.

We thank the reviewer for raising this point. We acknowledge that the statistical power to detect a significant difference in pre-miR-485 levels in Figure 2B may be limited due to the relatively small sample size and the inherent variability in hiPSC-derived neuronal cultures. However, it is important to emphasize that the functional impact of miRNAs is primarily mediated by their mature transcript forms. Our miRNA-seq data (Supplementary Table 2 and Figure S2) did not show significant alterations in the levels of mature miR-485-5p or miR-485-3p. This finding aligns with the reported expression pattern of miR-485 in hiPSC-derived neurons, where relatively low levels are observed in early neuronal development, with increased expression occurring in older, more mature neurons (Soutschek et al. 2023; https://ethz-ins.org/igNeuronsTimeCourse/ database from the Institute of Neurogenomics, ETH Zurich). This database provides a valuable resource for examining gene expression dynamics during human neuronal differentiation. Given that our hiPSC-derived neurons were analyzed at a relatively early developmental stage (DIV8 for these experiments), it is likely that miR-485 expression had not yet reached levels sufficient to reveal significant differences. While we acknowledge the potential limitation in statistical power for detecting subtle changes in pre-miR-485 levels, the combined evidence suggests that miR-485 may not be a significant contributor to the observed phenotypes at this developmental stage.

A paragraph has been added in the corresponding Results section to address this issue.

(5) There are a few situations where the authors could help out the reader a little bit by providing more labels on the figures directly. For example: in Figure 2, there are expression levels, over-expression, and inhibition of miRNA but the X-axis is named with similar labels for the miRNAs in question for each of these distinct experiments. If the authors want to help the reader, they may consider labeling these panels with a descriptive title to reflect the experiment being done or use more descriptive terms in the X-axis panels. Again, this is minor. Similarly, in Figure 5, it might be helpful for the authors to help out the reader again with more labels on the panels, such as in Figures 5B, 5C, and 5D. Would they consider labeling these panels, HPC, PFC, SSC with the brain location as they did in Figure 4?

We thank the reviewer for these helpful suggestions to improve the clarity of our figures. We have implemented the proposed changes. In Figure 2C-E, we have added specific titles to the panels to clearly distinguish between the different experimental conditions such as miRNA overexpression and inhibition. Similarly, in Figure 5, we labeled panels 5B, 5C, and 5D with the brain regions analyzed (HPC, PFC, SSC) to match the labeling used in Figure 4. We believe these revisions enhance the readability and overall interpretability of the figures, making it easier for readers to follow the experiments and results.

(6) Figure 3: There is some overshoot of the data in EMC10 homozygous null, in panel 3E, and also, overshoot of the het in panel 3H. Would there be value in the authors commenting on the potential basis for this in the discussion? Some issues are minor, such as the lack of electrophysiological analysis of circuits in vivo or in ex vivo slices that may further support the proposed rescue.

The reviewer correctly highlights the observation in Figures 3E and 3H, where the number of branch points in the Q6/EMC10HOM line exceeds wildtype levels and the calcium response in the Q6/EMC10HET and Q6/EMC10HOM lines surpasses that of the control. This overshoot is indeed intriguing and warrants discussion. EMC10 is part of the ER Membrane Complex (EMC), which plays a critical role in the proper folding and localization of various membrane proteins, including neurotransmitter receptors and ion channels such as voltage-gated calcium channels (Chitwood et al., 2018; Shurtleff et al., 2018; Chitwood and Hegde, 2019). In the context of the 22q11.2 deletion, EMC10 dysregulation may disrupt the proper localization of these proteins at the synapse, affecting both dendritic morphology and calcium signaling. The precise basis of this overshoot remains unclear. The overshoot may result from a dosage-sensitive inhibitory effect of Emc10, where both reduced and increased expression alter normal neuronal processes, with excessive responses potentially triggered upon gene restoration by the mutant system’s adaptation to dysfunction, leading to altered receptor sensitivity or signaling dynamics. This underscores the critical importance of precise Emc10 expression for proper neuronal development and function, in line with previous findings suggesting that EMC10 plays an auxiliary or modulatory role in EMC function. A short comment on the potential basis for this overshoot has been added in the corresponding Results section of the manuscript. Regardless of the underlying mechanisms, these findings emphasize the importance of precise titration of ASO constructs, rigorous gene dosage controls, and thorough analysis of context-specific responses to ensure both efficacy and safety in clinical applications.

We also agree with the reviewer that electrophysiological studies, particularly in the 22q11.2 deletion mouse model, would provide valuable insights into the impact of EMC10 modulation by ASOs on neuronal activity and circuit function at the in vivo and ex vivo levels. Incorporating such experiments into future studies will allow us to assess synaptic transmission and plasticity, contributing to a more comprehensive understanding of the therapeutic potential of ASO-mediated EMC10 modulation in 22q11.2DS.

(7) Did the authors take out the behavior studies further than 9 weeks? Would the authors consider commenting on what they speculate might be the duration of the treatment effect? For both mice and definitely humans.

We thank the reviewer for raising the important question regarding the duration of the ASO treatment effect, which is crucial for translating our findings into clinically relevant therapies. While behavioral studies beyond 9 weeks were not conducted in this study, our in vivo experiments and findings from prior publications (detailed below) enable an informed speculative assessment.

We utilized 2'-O-methoxyethyl (2'-MOE) modified ASOs, known for their enhanced binding affinity, nuclease resistance, and increased metabolic stability. In our in vivo post-injection screening of ASOs (Figure S13C), we predicted that Emc10 expression levels return to normal WT levels (~T100%) approximately 26 weeks post-treatment in Emc10ASO (#1466182) treated mice. This prediction is supported by our Emc10 expression profiles across various brain regions, which demonstrate robust repression of Emc10 lasting up to 10 weeks post-administration (Figure 6D-F). While these findings suggest that the treatment effect in our model could extend significantly beyond 10 weeks following a single ASO injection, further empirical validation is required through extended follow-up studies. Encouragingly, long-term effects of 2'-MOE ASOs have been observed in other neurological disorders (Kordasiewicz et al., 2012; Scoles et al., 2017; Finkel et al., 2017; Darras et al., 2019). However, factors such as ASO distribution, target cell turnover, and disease-specific pathophysiology could influence the duration of the effect. To address these uncertainties, we have added a paragraph in the Discussion section emphasizing the need for additional studies, including extended follow-up periods and eventual clinical trials, to determine the specific duration of effect for our Emc10ASO constructs in treating 22q11.2DS.

Reviewer #2 (Recommendations For The Authors):

(1) It is acknowledged that the iPSC-derived cells in Figure 1 are no longer progenitors, but differentiation markers for astrocytes and glia are also needed in Figure 1b to establish that equal rates of differentiation have occurred across genotypes.

We thank the reviewer for raising this important point about ensuring equal rates of differentiation across genotypes. As the reviewer notes, we employed a well-established protocol for directed differentiation of hiPSCs into cortical neurons using a combination of small molecule inhibitors, as previously described by Qi et al. (2017). This protocol has been extensively validated and is known to robustly generate cortical neurons while actively suppressing glial differentiation, as evidenced by the lack of upregulation of glial markers such as GFAP, AQP4, or OLIG2 in the original study. Given the established neuronal specificity of this protocol and our focus on neuronal phenotypes, we prioritized the confirmation of successful neuronal differentiation using the established neuronal markers TUJ1 and TBR1. Therefore, additional markers for astrocytes and glia are not included in this figure, as we did not expect significant glial differentiation under these conditions. A sentence has been added in the corresponding Results section to address this issue.

(2) For the RNA-seq experiments outlined in Figures 3J and K, a more comprehensive analysis is needed of the genes disrupted in the parental Q6 line relative to the het and homo lines. What percent are rescued, unaffected, vs uniquely disrupted?

Reduction in EMC10 levels is not expected to directly affect transcription or broadly reorganize the gene expression profile of the Q6/Q5 NGN2-iN lines. Our transcriptional profiling was not designed to assess the direct impact of EMC10 deficiency on gene expression but rather to measure the cellular pathways affected by reduced EMC10 in the patient Q6 line. We identified genes differentially expressed between the Q6 (patient) and Q5 (control) lines, whose expression differences were either abolished or significantly attenuated ("rescued") in the Q6/EMC10HET or Q6/EMC10HOM lines. In the Q6/EMC10HET line, 237 DEGs (6%) were rescued, while in the Q6/EMC10HOM line, 382 DEGs (11%) were rescued. Importantly, further analysis revealed 103 shared rescued DEGs in these lines, which was statistically significant (enrichment factor = 1.7; p < 0.0001, based on a hypergeometric test). We added a new figure panel (Figure 3L) to visualize the significant overlap of rescued DEGs from the Q6/EMC10HET and Q6/EMC10HOM lines. This overlap suggests these genes play a critical role in biological pathways impacted by EMC10 levels, particularly in nervous system development, as indicated by our functional annotation analysis. We also performed protein-protein interaction (PPI) network analysis to explore the functional relationships among these 103 shared DEGs (Figure S8). Future studies will further investigate these gene sets to gain deeper insights into the molecular mechanisms underlying 22q11.2DS and the role of EMC10.

(3) The authors claim that 50% EMC10 loss in adult mice is safe and should be toned down. EMC10 knockout mice have motor, anxiety, and social phenotypes. It would be unique amongst highly dosage-sensitive genes (MeCP2, CDKL5, TCF4, FMR1, etc.) for there to only be a neurodevelopmental component. In all those cases, and others, the effects of over and under-expression are reversible into adulthood. Establishing the range in adults is critical to establishing therapeutic utility. Absent a detailed examination of non-cognitive phenotypes, this claim cannot be made.

The reviewer raises an important point about the potential effects of EMC10 reduction in adult mice and the need to establish a safe therapeutic window by evaluating both cognitive and non-cognitive phenotypes. We agree that such a comprehensive evaluation is critical for assessing the safety and translational potential of Emc10-targeting therapies. While the International Mouse Genotyping Consortium reported motor and anxiety phenotypes in homozygous Emc10 knockout mice, these data are unpublished and based on a relatively small number of animals. Furthermore, in our previous work (Diamantopoulou et al., 2017), we demonstrated that complete Emc10 loss does not impair cognition or social behavior, as assessed by prepulse inhibition (PPI), working memory (WM), and social memory (SM) assays (see Figure 3A-D; Diamantopoulou et al., 2017). Additionally, heterozygous Emc10 mice, which exhibit a ~50% reduction in Emc10 expression similar to that achieved with our ASO treatment, showed no evidence of motor deficits or anxiety-like behavior. Specifically, Emc10+/- mice displayed locomotor activity comparable to WT mice in the open field (OF) test (Figure S4A, Diamantopoulou et al., 2017). Moreover, genetic normalization of Emc10 expression in Df(16)A+/- mice demonstrated no signs of anxiety-like behavior, as assessed by the OF test (Figure S4A) and elevated plus maze (EPM) (Figure S4B; Diamantopoulou et al., 2017). To further support these findings, we have added new data to the current manuscript (see Figure S10J) showing that TAM treatment-mediated restoration of Emc10 levels in the brain of adult Df(16)A+/- mice did not affect the time that mutant mice spent in the center area of the OF (Fig. S10J), suggesting that Emc10 reduction does not influence anxiety-related behavior. These results suggest that a 50% reduction in EMC10 expression is unlikely to result in motor or anxiety-like phenotypes in adult mice. Finally, as noted in the manuscript, in addition to prior findings from animal models, a substantial number of relatively rare LoF variants or potentially damaging missense variants have been identified in the human EMC10 gene among likely healthy individuals in gnomAD, a database largely devoid of individuals known to be affected by severe neurodevelopmental disorders (NDDs).

Nevertheless, the Discussion has been revised to underscore the importance of establishing a more detailed safety profile, including non-cognitive phenotypes, to fully validate the therapeutic potential of Emc10-targeting approaches. It also highlights the need for future studies to expand on these evaluations, addressing this critical aspect and laying a stronger foundation for advancing these findings into clinical drug development

(4) Supplemental Figure 10: The protein validation of Emc10 knockout following tamoxifen injection needs to be validated in all brain regions, not just the PFC. This is particularly important as the rest of the paper focuses on HPC-mediated phenotypes.

First, we want to emphasize that we conducted both qRT-PCR and WB assays on the same animal cohort, specifically examining the left and right hippocampus following ASO injection (see Figure S11C and D). This approach is crucial, given the central role of hippocampus in the phenotypes investigated in our ASO-mediated Emc10 knockdown experiments.

The reviewer raises an important point regarding the validation of EMC10 reduction at the protein level across all relevant brain regions using the Emc10 conditional knockout strain. We agree that such validation would ideally confirm the efficacy of our tamoxifen-induced knockout model comprehensively. However, we hope the reviewer appreciates that obtaining sufficient high-quality protein for WB analysis from smaller brain regions like the hippocampus poses a significant technical challenge. This difficulty is further compounded by the need to reserve the same samples for qRT-PCR to ensure consistency between mRNA and protein measurements. Importantly, our data from ASO-mediated Emc10 knockdown experiments (Figures S11C-D) demonstrate a clear and consistent correlation between reductions in Emc10 mRNA and protein levels in both the left and right hippocampus. Furthermore, in our constitutive Emc10-knockout mouse model (Diamantopoulou et al., 2017; see Figure S1A-B), we observed a strong agreement between mRNA and protein levels, supporting the reliability of mRNA data as a proxy for EMC10 protein levels in our experiments. Importantly, in all instances where we performed parallel protein and RNA measurements in human cell lines, there was excellent concordance between the datasets. Thus, while we acknowledge the limitations of relying primarily on mRNA data, we are confident that the Emc10 mRNA expression data in Figure S10 accurately reflect protein-level changes across brain regions in our conditional knockout model. To address this concern more fully in the future, we are working to refine antibody detection and optimize our protein extraction protocols to enable more routine and precise protein-level validation across smaller brain regions. We appreciate the reviewer’s feedback and will continue to refine our methodologies to strengthen the robustness of our findings.

(5) Figure 3: 1 way ANOVA would be more appropriate to analyze the data in B-G than t-tests.

We appreciate the suggestion of the reviewer. As mentioned above, we carefully selected statistical tests appropriate for each analysis. For Figure 3B-G, we chose to use pairwise t-tests to address specific hypotheses regarding the disease phenotype and rescue effects. This approach is consistent with prior experimental studies in the field, including our own (e.g., Xu et al., 2013; Figure 7H-I). Importantly, most of our t-tests yielded highly significant results (p < 0.001 or p < 0.01), reinforcing the robustness of our findings.

(6) Figure 5-6: Protein data is needed to complement the mRNA knockdown data.

We agree with the reviewer on the importance of protein-level validation to complement the mRNA knockdown data. As mentioned in our response to Reviewer’s Comment (4), in all instances where we performed parallel protein and RNA measurements, either in mouse brain or human cell lines, we observed excellent concordance between the datasets. This supports the reliability of our mRNA data as a proxy for protein changes. Nevertheless, we acknowledge the value of including protein validation in future experiments and will consider incorporating it to further strengthen our findings.

(7) The use of additional phenotypic measures is applauded in Figure 6, however, to appropriately interpret the data more is needed. Shao et al 2021 (Figure S9) show data from the International Mouse Genotyping Consortium claiming EMC10 KO mice have gait, activity, and anxiety phenotypes. All of these parameters could impact the SM assay and the y-maze assay. Changes in SM interaction time could be linked to anxiety or motor impairments, but interpreted as cognitive deficits because these symptoms were not assessed. At a minimum, discussion is needed about this limitation, as well as the inclusion of distance explored in the SM and Y-maze assays.

We thank the reviewer for their insightful comment regarding the potential influence of locomotor, gait, or anxiety phenotypes on the observed deficits in the SM and Y-maze assays. The behavioral phenotypes reported for Emc10 knockout mice by the International Mouse Genotyping Consortium (https://www.mousephenotype.org/data/genes/MGI:1916933) were limited to homozygous female mice and based on a small sample size (4–6 females) compared to a larger WT control group. Moreover, these data are unpublished and thus challenging to evaluate fully. Importantly, no abnormal behaviors were reported for Emc10 heterozygous knockout mice in these datasets. Additionally, the claim by Shao et al. (2021) regarding cognitive impairments in Emc10 knockout mice based on our previous work (Diamantopoulou et al., 2017) is inaccurate.

Our analysis of both the constitutive Emc10 knockout model (Diamantopoulou et al., 2017) and the current conditional Emc10 heterozygous knockout model consistently demonstrates that Emc10 reduction does not affect locomotor activity or anxiety-like behavior. In our earlier characterization of constitutive heterozygous Emc10 knockout mice (Emc10+/-), we observed no signs of anxiety-like behavior or motor impairments in OF assays (see Figure 2A-B and Figure S4A, Diamantopoulou et al., 2017). Similarly, results from Df(16)A+/- mice with genetically normalized Emc10 expression [Df(16)A+/-; Emc10+/-] also showed no indications of anxiety-like behavior or locomotor changes in the OF and EPM assays (see Figure S4A-B, Diamantopoulou et al., 2017). Consistent with these findings, our current data from Df(16)A+/- mice with conditional Emc10 reduction in the brain show no significant differences in locomotor activity and anxiety-related measures as assessed by OF assays (Figure S10J). Furthermore, total arm entries in Y-maze assays conducted in Df(16)A+/- mice treated with Emc10 ASOs were comparable to controls (Figures S14C and G-H), providing additional support for the conclusion that locomotor activity remains unaffected in these models.

We further appreciate the reviewer’s suggestion that changes in social interaction time during the SM assay could be influenced by anxiety or motor impairments. However, we consider this scenario unlikely in our model. Interaction times during the first trial of the SM assay, which measures general social interest, are comparable between Df(16)A+/- mice with reduced Emc10 expression (either genetically or through ASO treatment) and WT controls (see Figures 4E, 5E, and S10G). These findings indicate that our mouse models do not exhibit inherent difficulties in initiating social interaction, as might be expected if motor impairments or heightened anxiety were present. Reduced social interaction is commonly used as a behavioral marker for anxiety in rodent studies (reviewed by Bailey and Crawley, Anxiety-Related Behaviors in Mice, 2009). “Anxious” mice typically exhibit decreased social interaction, spending less time engaging with other mice compared to non-anxious counterparts. However, the specific deficit we observe in the second trial of the SM assay—when mice are reintroduced to a familiar juvenile—is indicative of impaired social recognition memory, as previously documented for Df(16)A+/- mice (Piskorowski et al., 2016; Donegan et al., 2020). This deficit is distinct from the general social avoidance typically associated with heightened anxiety.

Based on our comprehensive assessment of locomotor activity, anxiety-related behaviors, and social interaction, we conclude that the observed rescue of social memory and spatial memory deficits in mice with reduced Emc10 expression is most likely due to improved cognitive function rather than alterations in motor or anxiety-related domains.

(8) For ASO optimization experiments, it is not sufficient to claim robust uptake. A quantitative measure is needed using the PO antibody showing what percentage of cells were positive for the ASO. Since the contention is that only Emc10 in excitatory neurons is important, it would be helpful if this also included a breakdown of ASO uptake in excitatory and inhibitory neurons and astrocytes.

We thank the reviewer for highlighting the importance of quantifying ASO uptake and assessing cell-type specificity. To address this, we have added new data to the panel, as shown in the high-magnification images in Figure S14A. These images provide evidence that a large majority of NeuN-positive neurons exhibit a strong ASO signal. Specifically, we observed widespread ASO uptake (green) that extensively colocalized with the neuronal marker NeuN (red) in both the hippocampus and prefrontal cortex. Quantitative analysis of this overlap indicates that over 97% of NeuN-positive neurons were ASO-positive, demonstrating efficient neuronal uptake. This robust neuronal uptake aligns with the significant normalization of Emc10 levels and the behavioral improvements observed in ASO-treated Df(16)A+/- mice, further supporting the functional efficacy of our approach in modulating Emc10 expression within the relevant neuronal populations. Overall, the observed ASO uptake in neurons, as demonstrated by IHC, combined with RNA assays and the behavioral improvements in treated mice, strongly supports the efficacy of our approach in targeting Emc10 expression in the intended neuronal populations.

(9) An interpretation is needed in Figure S3 as to why ~50% of the pathways increased are also present on the decreased list. Ie. G1/transition, viral reproductive process, pos regulator of cell stress, etc. 4/10 GO terms are present in both increased and decreased groups in A and 7/10 in B.

We thank the reviewer for pointing out the overlap between pathways enriched in both the upregulated and downregulated miRNA groups in Figure S3. This overlap likely reflects the complex nature of miRNA regulation, where individual miRNAs can target multiple genes within a pathway, and single genes can be regulated by multiple miRNAs, sometimes with opposing effects (reviewed in Bartel, 2009; Bartel, 2018). For example, in the “G1/S transition” pathway, upregulated miRNAs such as miR-92a-3p, miR-92b-3p, and miR-34a-5p may promote the transition by targeting cell cycle regulators like FBXW7, CDKN1C, and CDK6 (Zhou et al., 2015; Zhao et al., 2021; Oda et al., 2024). Conversely, downregulated miRNAs such as miR-143-3p and miR-200b are known to suppress the transition by targeting genes such as HK2 and GATA-4 (Zhou et al., 2015; Yao et al., 2013). Our analysis identified overlapping predicted target genes for both upregulated and downregulated miRNAs, supporting the notion that many genes are subject to complex regulation by multiple miRNAs with potentially synergistic or antagonistic effects. Thus, the enrichment of certain GO terms in both groups likely reflects this intricate interplay of miRNA-mediated gene regulation. Future investigations focusing on specific miRNA-target interactions within these pathways will be critical to fully elucidate the underlying mechanisms and better understand the functional consequences of these opposing regulatory effects.

Minor Concerns:

(1) Define SM before using it.

We have defined the SM assay in the main text upon its first mention, where we describe the assay and its relevance to cognitive function (see page 11 of the revised manuscript).

(2) Statistics have been run in Figure S2, but not presented. The text only states that the differences between groups are significant. Please add in.

We have revised the legend of Figure S2 to include the specific statistical test used (students t-tests) and the corresponding p-values.

(3) The switch from ASO1 to ASO2 between Figures 5 and 6 needs more discussion. Why were new ASOs generated when ASO1 worked?

We thank the reviewer for their question regarding the transition from Emc10ASO1 to Emc10ASO2 between Figure 4 and Figures 5-6. Emc10ASO1 served as our initial proof-of-concept ASO construct, successfully demonstrating the feasibility of inhibiting Emc10 mRNA expression and providing evidence for behavioral rescue in our mouse model. As outlined in the manuscript, Emc10ASO2 targets a different region of the Emc10 transcript (intron 1, Figure 5A) compared to Emc10ASO1 (intron 2, Figure 4A). This distinction provides an additional layer of validation for our targeting strategy and ensures specificity in modulating Emc10 expression. In addition, Emc10ASO1 exhibited limited distribution in the brain, primarily targeting the hippocampus with weaker inhibition of Emc10 in other regions such as the cortex (Figure 4C, right panel). Emc10ASO2 overcame this limitation and achieve broader brain distribution, as demonstrated by the qRT-PCR data in Figure 5C. Given that 22q11.2DS can affect multiple brain regions and cognitive domains beyond the hippocampus, achieving broader distribution of the ASO is critical for a more comprehensive assessment of therapeutic potential.

(4) Page 3: Define "LoF"

We have defined Loss-of-Function (LoF) in the main text where it is first mentioned in the Introduction, where we discuss the potential of using LoF mutations to devise therapeutic interventions (see page 3 of the revised manuscript).

References

Bailey and Crawley, Anxiety-Related Behaviors in Mice, In: Methods of Behavior Analysis in Neuroscience. 2nd edition. Boca Raton (FL): CRC Press/Taylor & Francis; Chapter 5, (2009).

Bartel, MicroRNAs: target recognition and regulatory functions, Cell 136(2):215-33, (2009).

Bartel, Metazoan MicroRNAs, Cell, 173(1):20-51, (2018).

Chitwood et al., EMC Is Required to Initiate Accurate Membrane Protein Topogenesis, Cell 175, 1507-1519 e1516, (2018).

Chitwood and Hegde, The Role of EMC during Membrane Protein Biogenesis, Trends Cell Biol. (5):371-384, (2019).

Darras et al., Nusinersen in later-onset spinal muscular atrophy: Long-term results from the phase 1/2 studies, Neurology 92(21), (2019).

Diamantopoulou et al., Loss-of-function mutation in Mirta22/Emc10 rescues specific schizophrenia-related phenotypes in a mouse model of the 22q11.2 deletion, Proc Natl Acad Sci U S A 114, E6127-E6136, (2017).

Donegan et al., Coding of social novelty in the hippocampal CA2 region and its disruption and rescue in a 22q11.2 microdeletion mouse model, Nat Neurosci 23, 1365-1375, (2020).

Finkel et al., Nusinersen versus Sham Control in Infantile-Onset Spinal Muscular Atrophy, N Engl J Med 377(18):1723-1732, (2017).

Kordasiewicz et al., Sustained therapeutic reversal of Huntington's disease by transient repression of huntingtin synthesis, Neuron 74(6):1031-44, (2012).

Oda et al., MicroRNA-34a-5p: A pivotal therapeutic target in gallbladder cancer, Mol Ther Oncol, 32(1):200765, (2024).

Piskorowski et al., Age-Dependent Specific Changes in Area CA2 of the Hippocampus and Social Memory Deficit in a Mouse Model of the 22q11.2 Deletion Syndrome. Neuron 89, 163-176, (2016).

Qi et al., Combined small-molecule inhibition accelerates the derivation of functional cortical neurons from human pluripotent stem cells. Nat Biotechnol 35, 154-163, (2017).

Scoles et al., Antisense oligonucleotide therapy for spinocerebellar ataxia type 2, Nature 44(7650):362-366, (2017).

Shao et al., A recurrent, homozygous EMC10 frameshift variant is associated with a syndrome of developmental delay with variable seizures and dysmorphic features, Genet Med 23, 1158-1162, (2021).

Shurtleff et al., The ER membrane protein complex interacts cotranslationally to enable biogenesis of multipass membrane proteins, Elife 7, (2018).

Soutschek et al., A human-specific microRNA controls the timing of excitatory synaptogenesis, bioRxiv, (2023).

Stark et al., Altered brain microRNA biogenesis contributes to phenotypic deficits in a 22q11-deletion mouse model. Nat Genet 40, 751-760, (2008).

Xu et al., Derepression of a neuronal inhibitor due to miRNA dysregulation in a schizophrenia-related microdeletion, Cell 152, 262-275, (2013).

Yao et al., miR-200b targets GATA-4 during cell growth and differentiation, RNA Biol.10(4):465-8, (2013).

Zhao et al., miR-92b-3p Regulates Cell Cycle and Apoptosis by Targeting CDKN1C, Thereby Affecting the Sensitivity of Colorectal Cancer Cells to Chemotherapeutic Drugs, Cancers 2;13(13):3323, (2021).

Zhou et al., miR-92a is upregulated in cervical cancer and promotes cell proliferation and invasion by targeting FBXW7, Biochem Biophys Res Commun 458(1):63-9, (2015).

Zhou et al., MicroRNA-143 acts as a tumor suppressor by targeting hexokinase 2 in human prostate cancer, Am J Cancer Res. 5(6):2056-6 (2015).

Version 2

Abstract

Abstract

Adults and children with the 22q11.2 Deletion Syndrome demonstrate cognitive, social and emotional impairments and high risk for schizophrenia. Work in mouse model of the 22q11.2 deletion provided compelling evidence for abnormal expression and processing of microRNAs. A major transcriptional effect of the microRNA dysregulation is up-regulation of Emc10, a component of the ER membrane complex, which promotes membrane insertion of a subset of polytopic and tail-anchored membrane proteins. We previously uncovered a key contribution of EMC10 in mediating the behavioral phenotypes observed in 22q11.2 deletion mouse models. Here we show that expression and processing of miRNAs is abnormal and EMC10 expression is elevated in neurons derived from 22q11.2 deletion carriers. Reduction of EMC10 levels restores defects in neurite outgrowth and calcium signaling in patient neurons. Furthermore, antisense oligonucleotide administration and normalization of Emc10 in the adult mouse brain not only alleviates cognitive deficits in social and spatial memory but remarkably sustains these improvements for over two months post injection, indicating its therapeutic potential. Broadly, our study integrates findings from both animal models and human neurons to elucidate the translational potential of modulating EMC10 levels and downstream targets as a specific venue to ameliorate disease progression in 22q11.2 Deletion Syndrome.

Introduction

Adults and children with the 22q11.2 Deletion Syndrome (22q11.2DS) demonstrate cognitive, social and emotional impairments (1–3). 22q11.2 deletions are also one of the strongest genetic risk factors for schizophrenia (SCZ) (4). There are currently no targeted therapies that address the underlying molecular mechanisms of 22q11.2DS. Previous work in a model of the 22q11.2 deletion, carrying a hemizygous 1.3-Mb deficiency on mouse chromosome 16 [Df(16)A], which is syntenic to the 1.5Mb 22q11.2 deletion [Df(16)A+/–mice] revealed a distinct behavioral and cognitive profile (5, 6). Molecular analysis of the Df(16)A+/– strain provided compelling evidence for abnormal processing of brain enriched microRNAs (miRNAs) (5, 7). The Df(16)A+/– related miRNA dysregulation is due to (i) hemizygosity of Dgcr8, a component of the “microprocessor” complex that is essential for miRNA production (8) and (ii) hemizygosity of miRNA genes residing within the deletion, including mir185. Reduction of miR-185 levels and to a lesser degree of miRNAs residing outside the deletion {such as miR-485 (7)} result in a de-repression of Mirta22/Emc10 gene (herein after referred to as Emc10), whose expression is under the repressive control of miRNAs (7). Indeed, comprehensive RNA profiling of Df(16)A+/– mice found that postnatal elevation in the expression of Emc10 gene represents a key transcriptional effect of the 22q11.2 deletion (7). Increased brain expression of Emc10 is recapitulated in Df(16)A+/− primary neurons (9) as well as in mouse models of the more common 3Mb 22q11.2 deletion (10). Other miRNA targets are dysregulated, but their levels of change are subtler and more variable. Emc10 encodes for a component of the ER membrane complex (EMC), which promotes membrane insertion and maturation of a subset of polytopic and tail-anchored membrane proteins including neurotransmitter receptors, channels, and transporters (11–18). Emc10 is a prenatally biased gene with high expression in embryonic life that gradually subsides after birth (7), a developmental pattern of expression conserved between mice, humans and nonhuman primates (19). Emc10 Loss-of-Function (LoF) mutation that leads to reduction of Emc10 levels rescues key cellular, cognitive and behavioral alterations in the Df(16)A+/– mice (19). However, whether similar beneficial effects could be achieved in human neurons and whether Emc10 normalization in the adult brain could reverse established cognitive deficits remained unknown.

Here we show that 22q11.2 deletion results in abnormal processing of miRNAs in human neurons and in turn drives misexpression of EMC10 as previously described in animal models (5). Human EMC10 expression is elevated in neurons derived from 22q11.2 deletion carriers and reversal of EMC10 expression leads to restoration of key morphological and functional alterations linked to 22q11.2 deletions, supporting normalization of EMC10 expression as a disease-modifying intervention. Toward this end, we also show that antisense oligonucleotide (ASO)-mediated Emc10 normalization in the adult mouse brain is effective at reversing cognitive alterations. Improvements in cognition are sustained for over two months post ASO administration, underscoring the potential of this approach for providing durable therapeutic benefits. The observations that ASO-mediated Emc10 reduction in adult mouse brain rescues cognitive deficits linked to 22q11.2 deletion strongly support a key contribution of Emc10 and Emc10-dependent membrane protein trafficking in mediating the effects of 22q11.2 deletions on cognitive function and pave the way towards translating these observations into potential disease-modifying therapeutic interventions.

Results

To investigate whether miRNA dysregulation and upregulation of EMC10 is also prominent in cortical neurons from patients carrying 22q11.2 deletions (Fig. 1A), we used hiPSC lines obtained from three independent 22q11.2DS/SCZ donors carrying a 3Mb deletion and diagnosed with SCZ, along with matched healthy controls (Supplementary Table 1) to ensure the robustness and generalizability of our findings across different genetic backgrounds. The first patient/control pair is derived from dizygotic twins discordant for the 22q11.2DS and SCZ [Q6 (22q11.2) and Q5 (Ctrl)] (Fig. S1A-D). The second patient/control pair is derived from siblings [Q1 (22q11.2) and Q2 (Ctrl)] while the third patient/control pair is a case and age/sex matched unrelated control pair from the NIMH Repository and Genomic Resource [QR27 (22q11.2) and QR20 (Ctrl)].

EMC10 is robustly upregulated in hiPSC-derived neurons from 22q11.2 deletion carriers.

(A) Schematic diagram depicting the human chromosome 22q11.2 region. Bright grey and red horizontal bars indicate the two most common hemizygous genomic deletions found in the 22q11.2 deletion syndrome. The location of the coding genes and non-coding RNAs (miRNAs, underlined) are shown for chromosome 22q11.2. The microprocessor DGCR8 (DiGeorge Syndrome Critical Region Gene 8) and mir185 are shown in bold. (B) Cortical marker TBR1 and pan-neuronal marker TUJ1 expression in cortical neurons as detected by immunocytochemistry at day 13 of differentiation. TBR1 (red), TUJ1 (green) and DAPI (blue) expression are shown. Scale bars: 100 μm. (C) Volcano plot showing differentially expressed miRNAs (DEmiRs) in cortical neurons at day 8 of differentiation. Significant DEmiRs (P value < 5%) are shown above red line; Q5 (Ctrl) n=3, Q6 (22q11.2) n=3. 153/133 miRNAs were significantly up- and downregulated in Q6 (22q11.2) hiPSC-derived cortical neurons, respectively. 22q11.2 deletion region residing miRNAs miR-185, miR-1286 and miR-1306 are highlighted. (D-F) Consistent upregulation of EMC10 mRNA in Q6 (22q11.2) line derived cortical neurons as assayed by qRT-PCR at (D) day 8 (P= 0.031; Q5: n=3, Q6: n=3), (E) day 20 (P= 0.0478; Q5: n=4, Q6: n=4) and (F) day 34 (P= 0.0358; Q5: n=3, Q6: n=3) of differentiation. (G) Western blot analysis showing upregulated EMC10 protein levels in Q6 (22q11.2) line derived cortical neurons at day 8 of differentiation. Tubulin was probed as a loading control. (H) Immunofluorescence images of NGN2 generated cells. Representative images of NGN2-iNs at DIV21 from Q5 (Ctrl) and Q6 (22q11.2) hiPSC lines identified via EGFP fluorescence and immunostained for neuronal dendrite marker MAP2 and the nuclear marker DAPI. Scale bars=100 µm. (I-K) qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21. (I) Upregulation of EMC10 mRNA in Q6 (22q11.2) line derived neurons compared to the healthy control line Q5 (P=0.0222). Q5 (Ctrl) n=4, Q6 (22q11.2) n=4. (J) Upregulation of EMC10 mRNA in Q1 (22q11.2) patient line compared to healthy control line Q2 (P=0.0441). Q2 (Ctrl) n=5 and Q1 (22q11.2) n=7. (K) Upregulation of EMC10 mRNA in QR27 (22q11.2) patient line compared to healthy control line QR20 (P=0.0414). QR20 (Ctrl) n=5 and QR27 (22q11.2) n=5. Data are presented as mean ±SEM, unpaired two-tailed t-test, *P<0.05, **P<0.01.

We examined whether 22q11.2 deletion results in abnormal processing of miRNAs in human neurons as we have previously described in animal models (5). We performed parallel small RNA/miRNA sequencing on DIV8 differentiated human cortical neurons from the sibling (Q5/Q6) pair derived using an approach that combines small-molecule inhibitors to repress SMAD and WNT signaling pathways to promote CNS fate (20). This protocol has been extensively validated and is known to robustly generate cortical neurons while actively suppressing glial differentiation. We confirmed the efficiency of differentiation using immunohistochemistry (IHC) and gene expression assays, which indicated the anticipated increase of TUJ1/TBR1 positive derived neurons and downregulation of embryonic stem cell marker OCT4 (Fig. 1B and Fig. S1E-G). We identified a number of mature miRNAs dysregulated in response to the 22q11.2 deletion (Fig. 1C, Fig. S2A, B, Supplementary Table S2). As a validation of our approach, we observed the expected downregulation of miRNAs miR-185, miR-1286 and miR-1306 that reside in the 22q11.2 locus (three other predicted 22q11.2 miRNAs, miR-649, miR-3618 and miR-4761 were not detected in DIV8 neurons) (Fig. S2A). Among miRNAs located outside the 22q11.2 region, we note downregulation of miRNAs such as miR-137, as well as miR-134 and several other members form the largest placental mammal-specific miRNA gene cluster miR379-410 (Fig. S2B) that have been previously implicated in neuronal development, differentiation and function (5, 21–29). We used the miRNA-target interaction network tool miRNet 2.0 (30) to perform target enrichment and network analysis for the dysregulated miRNAs and performed GO term enrichment analysis on this target interaction network. Affected biological processes were prominently centered on cell division and intracellular protein transport (Fig. S3A) whereas cellular components were associated with the nucleus and the perinuclear region (endoplasmic reticulum and Golgi apparatus) of the cytoplasm (Fig. S3B).

In addition to 22q11.2 deletion region miRNAs, lower abundance of miRNAs in cases is likely due to haploinsufficiency of the DGCR8 gene and is expected to result in upregulation of target genes. To identify candidate miRNA target genes we performed an unbiased evaluation of the transcriptional responses using bulk RNA sequencing on RNA collected from DIV8 differentiated cortical neurons derived from the patient (Q6) and the corresponding healthy dizygotic twin (Q5) line (Fig. S4A). We observed the expected downregulation of genes within the 22q11.2 locus in patient neurons (Supplementary Table 3). Further, RNA and protein expression characterization confirmed the reductions in the abundance of the 22q11.2 locus residing genes DGCR8 and RANBP1 (Fig. S4B, C). Among the differentially expressed genes (DEGs) 2094 were downregulated and 1937 were upregulated. As expected EMC10 expression was elevated in patient neurons while expression of other EMC subunits (EMC1-4, EMC6-9) detected in our DIV8 sequencing data did not show significant differences. GO term enrichment analysis on downregulated DEGs identified significantly altered biological processes centered on neurogenesis, neuronal development, and differentiation (Fig. S4D). Among the upregulated DEGs, the GO terms enriched were related to neuronal development as well as neuronal cilia assembly and structure (Fig. S4E).

Intersection of predicted targets of downregulated miRNAs and up-regulated DEGs identified 774 predicted targets of downregulated miRNAs (Fig. S4F, Supplementary Table 4) including EMC10. Notably, functional annotation revealed that predicted targets of downregulated miRNAs include genes that modulate neuronal development and are associated with GO terms such as endoplasmic reticulum and endomembrane system of neurons (Fig. S4G, H).

qRT-PCR assays confirmed a robust and significant upregulation of EMC10 levels in RNA extracted from cortical neurons derived from hiPSCs of the Q5/Q6 pair through SMAD/WNT signaling inhibition, at three distinct stages of in vitro maturation (Fig. 1D-F). Additionally, this upregulation was confirmed in protein extracts from cortical neurons at day 8 of differentiation (Fig. 1G). To examine whether transcriptional EMC10 upregulation is independent of the neuronal derivation method, we generated neurons via inducible expression of Neurogenin-2 (NGN2), a widely used protocol that generates a robust population of excitatory neurons (NGN2-iNs) within 3 weeks (31–34). MAP2 staining was used to demonstrate the successful neuronal differentiation of the hiPSC lines (Fig. 1H). qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21 confirmed transcriptional EMC10 upregulation in three independent pairs of patient and sex-/age matched healthy control lines (Fig. 1I-K). Taken together our results highlight a reproducible and robust upregulation of EMC10 in neurons derived from patients with 22q11.2 deletions, which is independent of the derivation method. It is noteworthy that in addition to monolayer cultures, EMC10 shows significant upregulation along the excitatory neuron lineage (radial glia, intermediate progenitors and excitatory neurons) but not in astrocytes, choroid or interneuron lineage cells, in patient forebrain organoids generated by the same hiPSCs lines used in the present study (35).

We have previously shown that upregulation of the murine orthologue of Emc10 is primarily due to downregulation of miR-185 and to a lesser degree of miR-485 (7). Both conserved and non-conserved binding sites at the 3’UTR of human EMC10 are predicted in silico for both miRNAs (Fig. S5A). Consistently, the observed upregulation in the levels of EMC10 gene is accompanied by a robust reciprocal decrease in the levels of the miRNA precursor of miR-185 at DIV8 as indicated both by our miRNA sequencing analysis (Fig. 2A, Supplementary Table 2) and follow up qRT-PCR assays (Fig. 2A). The miRNA precursor of miR-485 exhibited a modest but non-significant reduction in abundance (Fig. 2B), consistent with our miRNA sequencing profile (Fig. S2B). This may be attributed to the early developmental stage of the neurons, as miR-485 expression increase during neuronal maturation (36, 37). Collectively, these findings demonstrate a strong inverse correlation between EMC10 upregulation and miR-185 downregulation, while suggesting that miR-485 may play a less prominent role at this early stage of neuronal development. Notably, overexpression of miR-185 and miR-485 using miRNA mimics in human cortical neurons at DIV10 resulted in reduction of EMC10 expression levels in both the healthy control (Q5, Fig. 2C) and patient line (Q6, Fig. 2D). Furthermore, inhibition of endogenous miR-185 and miR-485 in the control line by using specific miRNA inhibitors increased EMC10 expression level (Fig. 2E) confirming the predicted conserved functionality of miR-185 and miR-485 miRNA binding sites in EMC10. It is worth noting that in addition to miR-185, non-conserved binding sites at the 3’UTR of human EMC10 are predicted in silico for two additional downregulated miRNAs residing within the 22q11.2 locus, miR-1286 and miR-1306 (Supplementary Table 5). The functionality of these miRNA binding sites in EMC10 and whether they contribute to the observed elevation of its expression in human neurons remains to be determined. Taken together our results confirm that miRNA dysregulation emerges in human neurons as a result of the 22q11.2 deletion and in turn drives misexpression of genes primarily involved in intracellular membrane and protein trafficking-related processes required for neuronal development and maturation. Among them, EMC10 represents a major downstream effector of the 22q11.2-linked miRNA dysregulation.

Altered miRNA expression in hiPSC-derived cortical neurons from 22q11.2 deletion carriers.

(A) Precursor miRNA expression level of miR-185 (P= 0.0038), predicted to target EMC10, are downregulated in Q6 (22q11.2) cortical neurons as assayed by qRT-PCR (Q5: n=3, Q6: n=3). (B) Precursor miRNA expression level of miR-485 (P= 0.0622), predicted to target EMC10, are downregulated in Q6 (22q11.2) cortical neurons as assayed by qRT-PCR (Q5: n=3, Q6: n=3). (C-E) miR-185 and miR-485 modulate EMC10 in human iPSC-derived cortical neurons. (C) qRT-PCR quantification shows reduced expression levels of EMC10 mRNA in Q5 (Ctrl) line derived cortical neurons at day 10 of differentiation transfected with miR-185 [one-way ANOVA, F (2, 6) = 6.079, P=0.0361; post hoc Bonferroni, P=0.0366] or miR-485 [post hoc Bonferroni, P=0.0464] mimics at day 8 of differentiation. Expression levels in miR-185 or miR-485 mimic-treated neurons were normalized to expression levels under scramble mimic controls treatment (n= 3, each treatment). (D) qRT-PCR quantification shows reduced expression levels of EMC10 mRNA in Q6 line-derived cortical neurons transfected with miR-185 [one-way ANOVA, F (3, 13) = 7.167, P=0.0044; post hoc Tukey, P=0.0345] or miR-485 [post hoc Tukey, P=0.0251] or a combination of both miRNA mimics [post hoc Tukey, P=0.0020]. Expression levels in miR-185, miR-485 or the combination of both mimic-treated neurons were normalized to expression levels under scramble mimic controls treatment. Ctrl mimic n=5, miR-185 mimic n=5, miR-485 mimic n=4 and miR185+miR485 mimics n=3. (E) qRT-PCR quantification shows increased expression levels of EMC10 mRNA in Q5 line derived cortical neurons transfected with miRNA inhibitors miR-185 [one-way ANOVA, F (2, 6) = 11.94, P=0.0081; post hoc Bonferroni, P=0.0057]or miR-485 [post hoc Bonferroni, P=0.0491] at day 8 of differentiation. Expression levels in miR-185 or miR-485 inhibitor-treated neurons were normalized to expression levels under scramble miRNA inhibitor controls treatment (n=3, each treatment). Data are presented as mean ±SEM, unpaired two-tailed t-test or one-way ANOVA as indicated, *P<0.05, **P<0.01.

To investigate the relevance of EMC10 de-repression in the development and function of patient neurons, we generated derivatives of the Q6 patient hiPSC line carrying either heterozygous (Q6/EMC10HET) or homozygous (Q6/EMC10HOM) EMC10 LoF mutations using standard CRISPR/Cas9 editing approaches (Fig. S6A). Mutations were confirmed by sequencing (Fig. S6A, lower panel) and karyotyping confirmed normal chromosome complement (Fig. S6B). We confirmed reduced expression levels of 22q11.2 gene RANBP1 by western blot in both derivative hiPSC lines (Fig. S6C) whereas stem-cell marker NANOG and OCT4 were equally expressed in all lines assayed by qRT-PCR (Fig. S6D, E). EMC10 mRNA and protein levels were reduced by ∼50% in the Q6/EMC10HET hiPSC line and abolished in the Q6/EMC10HOM line (Fig. S6F, G). It is noteworthy that we did not observe an upregulation of EMC10 mRNA levels in the Q6 hiPSC lines (Fig. S6F), a finding likely attributed to the general low expression level of miR-185 and miR-485 in hiPSCs (38). Indeed both miRNAs are developmentally regulated and show increased expression levels during neuronal development (https://ethz-ins.org/igNeuronsTimeCourse/) (37). Additional characterization of hiPSC-derived NGN2-iNs (Fig. 3A), conclusively demonstrated a reduction (Q6/EMC10HET) or elimination (Q6/EMC10HOM) of EMC10 mRNA (Fig. 3B). Expression assays of a panel of cell type-specific markers did not reveal significant differences between NGN2-iNs from the Q6 patient line and both derivative lines, indicating that gene editing has no adverse effect on neuronal differentiation (Fig. S6H).

Reduction of EMC10 levels restores defects in neurite outgrowth and calcium signaling in neurons from 22q11.2 deletion carriers.

(A) Representative images of NGN2-iNs at DIV21 from Q6/EMC10HET and Q6/EMC10HOM hiPSC lines identified via EGFP fluorescence and immunostained for neuronal dendrite marker MAP2 and the nuclear marker DAPI. Scale bars=50 µm. (B) qRT-PCR assay of EMC10 mRNA expression level in NGN2-iNs at DIV21. EMC10 expression is normalized to near WT levels level in Q6/EMC10HET line (P=0.166) and abolished in Q6/EMC10HOM line (P<0.0001). Q5 (Ctrl) n=4, Q6 (22q11.2/SCZ) n=4, Q6/EMC10HET n=5 and Q6/EMC10HOM n=7. (C-G) Neuronal morphology analysis in Q5, Q6 (22q11.2), Q6/EMC10HET and Q6/EMC10HOM neurons. (C) Representative images of traced neurons. (D) Total neuronal length is reduced in Q6 line (P=0.0044) and restored in Q6/ EMC10HET (P=0.0253) and Q6/EMC10HOM line (P=0.0001). (E) Reduction in number of branch points/cell in Q6 (P=0.0195) is restored in the Q6/EMC10HET (P=0.0134) and Q6/EMC10HOM lines (P<0.0001). (F) Reduction in the total number of dendrites/cells in Q6 (P=0.0202) is reversed in the Q6/EMC10HET (P=0.0166) and Q6/EMC10HOM lines (P=0.0005) (G) The number of primary dendrites per cell is unchanged. Q5 (Ctrl) n=14, Q6 (22q11.2) n=16, Q6/EMC10HET n=19 and Q6/EMC10HOM n=16 neuronal cells. (H-I) Defects in cytoplasmic calcium signaling in Q6 (22q11.2) neurons are reversed in Q6/EMC10HET and Q6/EMC10HOM lines. (H) Changes in Fluo4-AM fluorescence signal intensity in response to 75mM KCl in Q5 (Ctrl), Q6 (22q11.2). Q6/EMC10HET and Q6/EMC10HOM hiPSC-derived neurons at DIV38. Q5 vs. Q6 (KS D=0.5405, P<0.0001), Q6 vs. Q6 EMC10HET (KS D=0.5556, P<0.0001) and Q6 vs. Q6 EMC10HOM (KS D=0.5676, P<0.0001). (I) Quantification of KCl-induced Fluo4 intensity peak amplitude (ΔF) demonstrates a reduction in Q6 line (P<0.0001) that is reversed in Q6/EMC10HET (P<0.0001) and Q6/EMC10HOM lines (P<0.0001). Q5 (Ctrl) n=70, Q6 (22q11.2/SCZ) n=37, Q6/EMC10HET n=82, Q6/EMC10HOM n=97 neuronal cells. (J) Heatmap (left) showing the expression of differentially regulated 237 genes in Q5 (Ctrl) and Q6 (22q11.2) that are normalized in the Q6/EMC10HET NGN2-iNs at DIV21 (n=3 per genotype). Gene Ontology (GO) biological process (BP) terms (right) associated with the up- and down-regulated genes normalized in the Q6/EMC10HET NGN2-iNs. (K) Heatmap (left) showing the expression of differentially regulated 382 genes in Q5 (Ctrl) and Q6 (22q11.2) that are normalized in the Q6/EMC10HOM NGN2-iNs at DIV21 (n=3 per genotype). Gene Ontology (GO) biological process (BP) terms (right) associated with the up- and down-regulated genes normalized in the Q6/EMC10HOM NGN2-iNs. (L) Intersection of rescued up- and down-regulated DEGs in the Q6/ EMC10HET and Q6/EMC10HOM lines: Venn diagram highlighting the 103 rescued DEGs (enrichment factor = 1.7; p < 0.0001, based on a hypergeometric test). Data are presented as mean ±SEM, unpaired two-tailed t-test or Kolmogorov–Smirnov test as indicated, *P<0.05, **P<0.01, ***P<0.001, ****P<0.0001.

Df(16)A+/− mice show impaired formation of dendrites in deep layer cortical neurons, which are faithfully recapitulated in primary neuronal cultures and are partially reversed by reduction of Emc10 levels (7). We asked whether impaired dendritic formation is also observed in human neuronal cultures from patients with 22q11.2 deletions and whether reduction of EMC10 levels could prevent such morphological alterations during neuronal maturation. We employed monolayer neuronal cultures of NGN2-iNs. Neuronal cells were fixed at DIV21 of differentiation, immuno-stained, traced and key indices of dendritic architecture were quantified (see Materials and Methods). Our analysis confirmed a reduced dendritic complexity in mutant neurons as reflected in total neuronal length, the number of branch points and the total number of dendrites per cell (Fig. 3C-F). The number of primary dendrites per cell were unchanged (Fig. 3G) in accordance with previous findings from the murine 22q11.2 deletion model where only subtle changes were detected in the number of primary neurons (7). Importantly, we found that reduction or elimination of EMC10 expression restored to WT levels neuronal length and branchpoints. Importantly, we found that reducing or eliminating EMC10 expression restored neuronal length and branch points to WT levels. Notably, the number of branch points in Q6/EMC10HOM neurons exceeded those in WT neurons (Fig. 3E) likely suggesting that reduced (or abolished) Emc10 expression can alter normal neurite growth, resulting in excessive responses, potentially triggered upon gene restoration by the mutant system’s adaptation to dysfunction, leading to altered receptor sensitivity or signaling dynamics. This highlights the critical importance of precise Emc10 expression for maintaining proper neuronal development and function.

Our previous evaluation of Ca2+ homeostasis perturbations caused by 22q11.2 deletions using Ca2+ imaging on primary neurons from Df(16)A+/– mice revealed a significantly lower amplitude of Ca2+ elevation following KCl evoked depolarization (9). This impairment was replicated in human cortical neurons from patients with 22q11.2 deletions (39) and shown to be partially restored by exogenous expression of DGCR8, indicating a potential role of miRNA dysregulation. Using the green-fluorescent calcium indicator Fluo-4 and time-lapse microscopy we confirmed a decrease in the amplitude of Ca2+ rise following KCl evoked depolarization, in patient (Q6) derived NGN2-iNs at DIV37/38 compared to the healthy twin (Q5) (Fig. S7A, B). We asked whether reduction of EMC10 levels could reverse such alterations. Notably, the observed defect in Ca2+ signaling were reversed in both Q6/EMC10HET and Q6/EMC10HOM NGN2-iNs as demonstrated by the increased amplitude of Ca2+ rise following depolarization (Fig. 3H, I). Interestingly, the amplitudes of Ca2+ rise in Q6/EMC10HET and Q6/EMC10HOM were slightly elevated compared to the WT control group, consistent with the effects on neurite outgrowth (Fig. 3H). The observation that reduction of EMC10 levels fully restores the Ca2+ signaling deficits observed in patient neurons suggests that miRNA-dependent elevation of EMC10 may interfere with one or more sources of intracellular Ca2+ and a wide range of calcium-dependent processes.

Differentially expressed genes are often organized into functional groups or pathways based on their known biological roles. We used transcriptional profiling as an indirect measure of cellular pathways affected by the reduction of EMC10 levels by identifying genes differentially expressed between the parental Q6/Q5 lines whose expression differences are abolished or nearly abolished (“rescued”) in either Q6/EMC10HET or Q6/EMC10HOM NGN2-iNs (Fig. 3J, K). In the Q6/EMC10HET line, 237 DEGs (6%) were rescued (111 downregulated and 126 upregulated), while in the Q6/EMC10HOM line, 382 DEGs (11%) were rescued (193 downregulated and 189 upregulated) (Supplementary Table 6). In both cases, functional annotation analysis indicated highest enrichment scores for terms related to nervous system development as well as an enrichment in GO terms relevant to neuronal generation and differentiation. Intersection of “rescued” genes in Q6/EMC10HET and Q6/EMC10HOM NGN2-iNs identified 103 shared DEGs (Fig. 3L and Supplementary Table 7). To assess the significance of the overlap between the rescued DEGs in Q6/EMC10HET and Q6/EMC10HOM NGN2-iNs, we performed a hypergeometric test, which calculates the probability of observing the degree of overlap between two gene groups under the null hypothesis that the overlap occurs by chance. Our analysis revealed a significant overlap (enrichment factor = 1.7; p < 0.0001), indicating that the overlap is much greater than expected by chance. These 103 shared DEGs likely play a key role in pathways influenced by EMC10 levels, particularly those involved in nervous system development, as suggested by our functional annotation analysis. To further investigate the functional relationships among these shared DEGs, we conducted a protein-protein interaction (PPI) network analysis. This analysis highlighted a functional cluster including 30 of these genes, such as the SCZ-linked genes PCDHA2 (40), RBFOX1 (41) and RGS4 (42, 43) (Fig. S8A), involved in nervous system development (Fig. S8B). It should be noted that the beneficial effect of elimination of EMC10 expression is consistent with previous findings that lack of EMC10 does not compromise EMC assembly (44) and implying an auxiliary or modulatory role of EMC10 in the EMC function.

Taken together, our analysis of neurons from 22q11.2 deletion carriers indicate that elevation of EMC10 expression disrupts their development and maturation in a way similar to observations in murine neurons, and support normalization of EMC10 expression as a disease-modifying intervention. While our previous work has shown that constitutive genetic reduction of Emc10 levels rescues key cognitive and behavioral alterations in the Df(16)A+/– mice, translating these observations into therapeutic interventions requires demonstration that it is the sustained elevation of EMC10 throughout the adult life that interferes with the underlying neural processes rather than an irreversible impact on brain maturation during early development. Toward this end, we first investigated whether restoration of Emc10 levels in the brain of adult (2– 4-month-old) Df(16)A+/– mice is effective at reversing cognitive alterations (5, 19). Specifically, we examined the effects of Emc10 reduction in adult brain on social memory (SM), a cognitive domain robustly and reproducibly affected in adult Df(16)A+/–mice (6, 19, 45). Notably, SM deficits are also present in juvenile Df(16)A+/– mice as early as postnatal day 22 (Fig. S9A, B), underscoring the severity of this phenotype, which emerges during early adulthood. In humans, SM, a key component of social cognition, involves encoding, storing, and retrieving information about social experiences, such as recognizing familiar individuals and recalling past interactions and emotions. Social cognition, the broader ability to perceive, interpret, and respond to social cues, is essential for forming relationships and understanding others’ behavior. Disruptions in SM can impair social cognition, contributing to the functional deficits commonly observed in schizophrenia (46). Deficits in social cognition are present in individuals with 22q11.2 deletions (47) and use of rodent tasks that evaluate SM can serve as a useful proxy of the human condition. Impaired SM in Df(16)A+/–mice is fully restored by constitutive genetic reduction of Emc10 levels (19).

To manipulate the expression of the Emc10 gene in adult Df(16)A+/– mice we used a Emc10 conditional ‘knockout-first’ design by conducting a Flp- and Cre-dependent genetic switch strategy (Fig. S10A). Parental Emc10+/- tm1a mice were crossed to a germline Flp mouse line that activates global Flp function and leads to the deletion of the frt-flanked sequence(s) in the offspring. The Emc10tm1c offspring from this cross carry a loxP flanked WT Emc10 allele and are essentially WT. To achieve temporal control of Emc10 expression we used an inducible UBC-Cre/ERT2 mouse line that activates global Cre function upon tamoxifen (TAM) treatment. This approach enables postnatal normalization of Emc10 expression at its endogenous locus preserving Emc10 expression within its physiological levels. We used UBC-Cre/ERT2 mice in crosses to generate compound Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/–mice. These mice have two WT Emc10 copies upregulated, as expected in the Df(16)A background, until TAM-induced Cre expression deletes the tagged Emc10 allele. We used oral gavage to deliver TAM and implement Cre-mediated Emc10 deletion during adulthood (postnatal day 56-70). Corn oil treatment served as a control. Behavioral analysis was performed on the following four groups: Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/+ mice treated with TAM (WT + TAM), Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/–mice treated with TAM (Df(16)A+/–+ TAM), Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/+ mice treated with corn oil vehicle (WT + oil), and Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/– mice treated with corn oil vehicle (Df(16)A+/– + oil). Investigation of the efficiency of Cre-mediated deletion in brain lysate preparations from prefrontal cortex (PFC) (Fig. S10B, C), hippocampus (HPC) (Fig. S10D, E) and cerebellum (CB) (Fig. S10F) confirmed that upon TAM treatment, Emc10 mRNA and protein levels were restored to near WT levels in the adult brain of Df(16)A+/– mice. As expected, Df(16)A+/– + oil mice showed impaired SM performance compared to WT + oil control littermates, which was fully rescued upon TAM treatment. Specifically, upon reintroduction of a familiar juvenile mouse Df(16)A+/–+ TAM mice showed a strong reduction in social interaction, indicative of intact SM, comparable to TAM-treated WT littermates and significantly different from Df(16)A+/– mice treated with corn oil (Fig. S10G). The intact SM of the Df(16)A+/– + TAM mice was further evident in analysis of difference score (Fig. S10H) compared to the corn oil-treated Df(16)A+/– mice. Interaction times during the first trial of the SM assay, which measures general social interest, were unaffected by TAM treatment. In contrast to SM, Df(16)A+/–mice hyperactivity in the open field was not affected upon TAM treatment consistent with our previous results from constitutive genetic reduction of Emc10 levels (Fig. S10I). Notably, TAM treatment did not alter the time spent in the center area of the open field, indicating an absence of changes in anxiety-related behavior. These findings demonstrate that restoring Emc10 levels in adult Df(16)A+/–mice can significantly improve cognitive deficits, underscoring a broad therapeutic window and establishing Emc10 as a promising target for postnatal interventions.

We explored the translation potential of this finding by employing transient injection of single stranded ASOs targeting the mouse gene as dictated by their demonstrated efficacy as a therapeutic modality in preclinical models (48–53) and clinical studies of neurodevelopmental disorders (NDDs) or neurodegenerative disorders (54–56). Over 300 chimeric 2′-O-methoxyethyl (2’MOE)/DNA gapmer ASOs were generated and screened for Emc10 mRNA reduction in 4T1 cells via electroporation (Fig. S11A). Lead ASOs were then confirmed in a dose-response assay (Fig. S11B), including the lead ASO (1081815, herein referred to as Emc10ASO1), which targets intron 2 of mouse Emc10 (Fig. 4A). Emc10ASO1 was selected for subsequent studies, as it was effective in lowering Emc10 expression both in vitro and in vivo. Specifically, following transient intracerebral ventricular (ICV) injection in the posterior ventricle of 8 weeks old WT mice, Emc10ASO1 effectively suppressed the levels of Emc10 mRNA (Fig. S11C) and protein (Fig. S11D) in both left and right HPC compared to WT mice treated with a control ASO (CtrlASO1) without complementarity in the mouse. Analysis of Gfap and Aif1 expression did not reveal any changes (Fig. S11E, F) suggesting lack of astroglial and microglial activation upon Emc10ASO1 injection. Emc10ASO1 injected mice showed normal gait and no signs of behavioral toxicity. IHC analysis using an antibody that selectively recognizes the phosphorothioate backbone verified a robust diffusion primarily in HPC and to a lesser degree in surrounding brain areas. Colocalization with the neuronal marker NeuN and glial fibrillary acidic protein (GFAP) confirmed accumulation in hippocampal neurons as well as GFAP-expressing astrocytes (Fig. 4B). Analysis of Df(16)A+/–mice treated by intraventricular injection at 8 weeks of age showed that Emc10ASO1 effectively lowered hippocampal Emc10 mRNA to nearly WT levels 3 weeks post-injection resulting in normalization of Emc10 expression (Fig. 4C, left panel). By contrast, consistent with the pattern of ASO distribution, we did not observe a significant reduction of Emc10 expression levels in the PFC of Df(16) +/– mice treated with Emc10ASO1 (Fig. 4C, right panel). In addition to targeted assays, we performed bulk RNA-sequencing analysis of CtrlASO1 and Emc10ASO1-treated Df(16)A+/– mice and WT littermates to evaluate the effect of Emc10ASO1 treatment on the hippocampal transcriptome profile. In the CtrlASO1 treated group (Fig. 4D, left panel), we observe the tripartite differential gene expression signature characteristic of Df(16)A+/–mice: upregulation of Emc10 and non-coding RNAs (pri-forms of miRNAs and long non-coding RNAs (Fig. 4D, left panel and inset, Log2Fold Change = 0.5) as well as the expected downregulation of genes included within the Df(16)A deficiency. In the Emc10ASO1 treated group (Fig. 4D, right panel and inset), Emc10 is no longer upregulated in Df(16)A+/– mice while non-coding RNAs remain upregulated, and genes included in the deficiency are robustly downregulated. Apart from Emc10, seven other genes (Mir9-3hg, Plxnd1, Cd68, Mir22hg, Gm28439, Adgre1, and Tnn) are significantly upregulated in Df(16)A+/– mice in the CtrlASO1 but not in the Emc10ASO1-treated group (Supplementary Table S8). We used the Bowtie mapping tool (57) to align short sequencing reads on both genomic and transcript sequence to assess whether these expression changes represent potential off-target effects of the Emc10ASO1 in the mouse transcriptome. Emc10ASO1 exclusively aligned with full complementarity to an intronic region in the Emc10 gene (Fig. 4A) providing additional support for high target specificity. The observed changes might represent downstream effects of Emc10 level reduction or reflect expression variability due to low expression levels of the upregulated genes.

ASO-mediated suppression of murine Emc10 in vivo.

(A) Mouse Emc10 gene map plot (ENSMUST00000118808) showing the Emc10ASO1 target site. (B) Mouse brains collected three weeks post ICV injection were stained with an ASO antibody (green), counterstained with neuronal marker NeuN (red) and nuclear stain Hoechst (blue). A robust and uniform ASO diffusion (top panel) is observed in the HPC. No signal is detected in saline injected mice (bottom panel). Overlap with NeuN (yellow, top-right panel) confirms presence in neuronal cells. Accumulation in glial cells, specifically GFAP labeled astrocytes is also observed (middle-right panel and inset; ASO in green, GFAP in red, and Hoechst in blue). Images are taken with 4x, 20x and 40x objectives. (C) qRT-PCR analysis shows Emc10ASO meditated normalization of Emc10 mRNA levels in the HPC of Df(16)A+/- mice (left panel). Significant upregulation of Emc10 mRNA expression levels is seen in CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (2, 29) = 11.65, P<0.001; post hoc Tukey, P= 0.001]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P= <0.001). CtrlASO-treated WT mice: n=9 (5 males, 4 females), CtrlASO-treated Df(16)A+/ mice: n=11 (5 males, 6 females), and Emc10ASO-treated Df(16)A+/ mice: n=12 (7 males, 5 females). qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+/- (right panel) compared to WT mice [one-way ANOVA, F (2, 16) = 4.253, P=0.0330; post hoc Tukey, P= 0.0385]. ASO injection does not normalize Emc10 expression levels in PFC of Df(16)A+/ mice injected with Emc10ASO (post hoc Tukey vs CtrlASO treated Df(16)A+/ mice, P= 0.7642). CtrlASO treated-WT male mice: n=6, CtrlASO treated-Df(16)A+/- male mice: n=5, and Emc10ASO treated-Df(16)A+/ male mice: n=8. (D) Volcano plots showing up-regulation of Emc10 expression in the CtrlASO treated-Df(16)A+/- compared to the CtrlASO-treated WT mice (left panel and inset) but not in the Emc10ASO treated-Df(16)A+/- compared to the Emc10ASO-treated WT mice (right panel and inset). The expected down-regulation of genes included in the Df(16)A+/– deletion (blue) and upregulation of non-coding RNAs (ncRNAs, red) is also evident in both panels. Downregulated genes from the 22q11.2 locus (Dgcr8, Ranbp1 and Tango2) as well as the upregulated ncRNA (miRNA-containing) gene Mirg, are highlighted. CtrlASO treated-Df(16)A+/–males: n=5, Emc10ASO treated-Df(16)A+/– males: n= 4, CtrlASO treated WT males: n=4, and Emc10ASO treated-WT males: n=4. (E-F) ASO mediated behavioral rescue of SM deficit in Df(16)A+/- mice. (E) CtrlASO-treated Df(16)A+/- mice show a robust SM deficit compared to CtrlASO-treated WT mice as indicated by the significant difference in trial 2 interaction time upon reintroduction of a familiar juvenile mouse [three-way ANOVA for Trial X Genotype X Treatment Interaction matching by trial: F (1, 38) = 9.393 P=0.0040; post hoc Tukey, P= 0.0012]. A reduction in trial 2 interaction time indicates rescue of the SM deficit in Emc10ASO-treated Df(16)A+/- compared to CtrlASO-treated Df(16)A+/- mice (post hoc Tukey, P= 0.0114). (F) A negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+/- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction: F (1, 38) =9.369, P=0.0040; post hoc Tukey, P= 0.0002]. Increase in the difference score of Df(16)A+/- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P= 0.0004). [CtrlASO-treated WT mice: n=9 (5 males, 4 females), CtrlASO-treated Df(16)A+/ mice: n=11 (5 males, 6 females), Emc10ASO treated WT mice: n=9 (5 males, 4 females), and Emc10ASO-treated Df(16)A+/ mice: n=13 (7 males, 6 females)]. Data are presented as mean ± SEM, *P < 0.05; **P < 0.01; ***P < 0.001.

Eight-week-old Df(16)A+/– mice and WT littermates were treated by ICV injection of Emc10ASO1 and CtrlASO1 and tested 3 weeks later in SM assays. Df(16)A+/– mice treated with CtrlASO1 showed the expected deficits in SM as reflected in the sustained high interaction time with the reintroduced familiar juvenile mouse. By contrast, Df(16)A+/– mice injected with Emc10ASO1 had significantly improved memory performance to levels indistinguishable from Emc10ASO1-treated WT littermates, consistent with improvement of function arising from adult restoration of Emc10 levels (Fig. 4E, F). Interaction times during the first trial of the SM assay, which measures general social interest, were unaffected by ASO treatment. Rescue was observed in both sexes and no significant differences were seen in treatment across sexes. In control experiments, we did not observe any effects of genotype or treatment upon reintroduction of a novel juvenile mouse in trial 2 (Fig. S12A, B), strongly indicating that SM deficits are not driven by a simple task fatigue. To evaluate the consistency and reproducibility of the ASO-mediated SM rescue we generated additional ASOs targeting different regions within the Emc10 transcript and screened them for Emc10 mRNA reduction in vivo (Fig. S13A-C). One of these ASOs (1466182, herein referred to as Emc10ASO2), which targets intron 1 of mouse Emc10 (Fig. 5A) was selected as the lead ASO candidate for the replication analysis based on its efficacy in reducing Emc10 levels, distribution pattern, as well as the lack of any signs of astroglial/microglial activation or behavioral toxicity (see Materials and Methods). IHC analysis showed robust ASO distribution in both the hippocampus (HPC, Fig. S14A, first and second panel) and prefrontal cortex (PFC, Fig. S14A, third and fourth panel), as well as diffusion into both neuronal and non-neuronal cells. This was indicated by colocalization with the neuronal marker NeuN and the glial marker GFAP (Fig. S14A, bottom panel). Higher-magnification (40x) analysis revealed extensive overlap between the ASO signal and NeuN staining, demonstrating that the vast majority (>97%) of neurons exhibited ASO uptake. qRT-PCR analysis of Df(16)A+/–mice treated by intraventricular injection at 8 weeks of age showed that Emc10ASO2 effectively lowered Emc10 mRNA to nearly WT levels 3 weeks post-injection resulting in normalization of Emc10 expression in the HPC (Fig. 5B), PFC (Fig. 5C) and somatosensory cortex (SSC) (Fig. 5D). To study the effects of Emc10ASO2-mediated Emc10 reduction on SM performance, 8-week-old Df(16)A+/- male mice and WT littermates were treated by ICV injection of Emc10ASO2 and CtrlASO2 and tested 3 weeks later. Df(16)A+/- mice injected with Emc10ASO2 had significantly improved SM performance to levels indistinguishable from Emc10ASO2-treated WT littermates (Fig. 5E, F).

Restoration of cognitive function in Df(16)A+/- mice using an independent Emc10 ASO.

(A) Mouse Emc10 gene map plot (ENSMUST00000118808) showing the Emc10ASO2 target site. (B) qRT-PCR analysis shows significant upregulation of Emc10 mRNA expression levels in the HPC of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (2, 19) = 20.92, P<0.0001; post hoc Tukey, P= 0.0018]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P= <0.0001). (C) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (2, 19) = 40.75, P=<0.0001; post hoc Tukey, P=<0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in the Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P=<0.0001). CtrlASO-treated WT male mice: n=6, CtrlASO-treated Df(16)A+/ male mice: n=8, and Emc10ASO-treated Df(16)A+/ mice: n=8. (D) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in Somatosensory Cortex (SSC) of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (2, 18) = 21.53, P=<0.0001; post hoc Tukey, P=<0.0026]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P=<0.0001). CtrlASO-treated WT mice: n=6, CtrlASO-treated Df(16)A+/ mice: n=7, and Emc10ASO-treated Df(16)A+/ mice: n=8. (E) CtrlASO-treated Df(16)A+/- mice show a robust SM deficit compared to CtrlASO-treated WT mice as indicated by the significant difference in trial 2 interaction time upon reintroduction of a familiar juvenile mouse [three-way ANOVA for Trial X Genotype X Treatment Interaction matching by trial: F (1, 26) = 35.74 P<0.0001; post hoc Tukey, P= 0.0004]. A reduction in trial 2 interaction time indicates rescue of the SM deficit in Emc10ASO-treated Df(16)A+/- compared to CtrlASO-treated Df(16)A+/- mice (post hoc Tukey, P= 0.0255). [CtrlASO-treated WT male mice: n=6, CtrlASO-treated Df(16)A+/ mice: n=8, Emc10ASO treated WT mice: n=8, and Emc10ASO-treated Df(16)A+/ mice: n=8. (F) A negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+/- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 26) = 35.74, P<0.0001; post hoc Tukey, P<0.0001. Increase in the difference score of Df(16)A+/- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P<0.0001). (G) Y-maze task displayed memory impairments in adult Df(16)A+/- mice. Impaired short-term spatial memory in Df(16)A+/- mice shown by the reduced amount of delayed alternations (%) after a delay of 1h (P=0.0015). WT mice: n=12, Df(16)A+/ mice: n=11. (H) Deficits in short-term spatial memory in the Y-maze task in CtrlASO-treated adult Df(16)A+/- mice can be rescued in Emc10ASO-treated Df(16)A+/- mice (one-way ANOVA, F(3, 25) = 6.727, P=0.0018, post hoc Tukey, P=0.0042) after 3-weeks of ASO-injection. CtrlASO-treated WT mice: n=8, Emc10ASO treated WT mice: n=8, CtrlASO-treated Df(16)A+/ mice: n=8 and Emc10ASO-treated Df(16)A+/ mice: n=5. (I) In a contextual fear memory assay, minimal freezing is observed on day 1 (left) with no significant changes across groups [two-way ANOVA for Genotype X Treatment interaction: F (1, 52) = 1.003, P=0.3211]. In the CtrlASO-treated group, Df(16)A+/- mice show the expected contextual fear memory deficit compared to WT mice (right panel) [one-way ANOVA, F (3, 52) = 3.524 P=0.0212; post hoc Tukey, 0.0384]. Emc10ASO treatment was not sufficient to fully rescue the learning deficit in Df(16)A+/– mice compared to CtrlASO-treated WT levels (post hoc Tukey, P= 0.1045). However, there is increased freezing in Df(16)A+/- mice injected with Emc10ASO-versus CtrlASO-treated group, which results in a non-significant difference in freezing between Emc10ASO-treated Df(16)A+/- and WT mice (two-way ANOVA for genotype x treatment F (1, 52) = 0.8524, P=0.3601; post hoc Tukey, P= 0.4676) indicating a partial rescue of the contextual fear memory deficit. CtrlASO-treated WT: n=13 (9 males, 4 females), Emc10ASO treated WT mice: n=14 (10 males, 4 females), CtrlASO-treated Df(16)A+/ mice: n=15 (11 males, 4 females and Emc10ASO-treated Df(16)A+/ mice: n=14 (10 males, 4 females). Unpaired students t-test, one-two- or three-way ANOVA as indicated. Data are presented as mean ± SEM, *P < 0.05, **P < 0.01, ***P < 0.01, ****P<0.0001.

In addition to SM deficits, mouse models of the 22q11.2 deletion show a spectrum of cognitive impairments in episodic and spatial memory as reflected, for example, in impaired performance in an Y-maze-based delayed alternations task that probes short-term spatial memory (58) and contextual fear conditioning a form of associative learning test used for studying episodic learning and spatial memory (5). We investigated the impact of ASO-mediated Emc10 reduction in the adult brain on both of these cognitive tasks. First, we confirmed that adult male Df(16)A+/– mice exhibit impaired short-term spatial memory during novelty exploration in a two-trial delayed alternation Y-maze task (Fig. 5G) as previously described for another mouse model of the 22q11.2 deletion (60). The total number of arm entries remained unchanged, indicating no alternations in locomotor activity (Fig. S14B). To determine whether reducing Emc10 expression in the brain via ASO treatment could rescue short-term spatial memory deficits, we tested a new cohort of Df(16)A+/–mice and WT littermates 3-weeks following ASO administration (Fig. 5H). ASO-treated Df(16)A+/– mice exhibited a significant improvement in delayed alternations compared to Df(16)A+/– mice treated with control ASO (Fig. 5H). Furthermore, no significant differences in total number of arm entries were observed between the groups (Fig. S14C). We confirmed the reduction of Emc10 levels in the ASO-treated animals through qRT-PCR assays of the HPC, PFC and SSC brain regions (Fig. S14D-F). In the contextual fear conditioning task, while ASO treatment was not sufficient to fully rescue the learning deficit in Df(16)A+/– mice to WT levels (Fig. 5I, right panel), there was a modest improvement in fear memory of ASO-treated Df(16)A+/– mice, since these mice did not differ significantly from the ASO-treated WT littermates. Interestingly, we have previously shown that genetic reduction of Emc10 levels in Df(16)A+/– mice resulted in only partial restoration of deficits in contextual fear memory (19). Thus our finding faithfully recapitulates results from our previous constitutive genetic rescue assays (19) and likely indicates a more limited role of Emc10 upregulation in the 22q11.2-linked fear memory deficits rather than requirement for additional treatment time or for earlier onset of Emc10 normalization.

The application of ASOs as a novel therapeutic strategy has seen a significant rise in recent years, in part due to their versatility in durably modifying RNA transcripts. Therefore, we investigated the longevity of ASO-mediated repression of Emc10 as an indicator of future therapeutic relevance for the treatment of 22q11.2DS. To this end, we conducted the SM and Y-maze assays on a new cohort of Df(16)A+/– mice at 3-4 weeks and 8-9 weeks post injection with Emc10ASO2 (Fig. 6A). We observed behavioral rescue in Emc10ASO2-treated Df(16)A+/– mice in the SM assay at three weeks (Fig. 6B, left panel) and in the Y-maze assay at four weeks post ASO-administration (Fig. 6C, left panel) in accordance to our previous findings (Fig. 5F, H). Remarkably, we replicated these results at 8-9 weeks post injection, demonstrating sustained behavioral rescue of SM (Fig. 6B, right panel) and spatial memory deficits (Fig. 6C, right panel). Importantly, locomotor activity remained unchanged in the Y-maze assays at both, four weeks (Fig. S14G) and nine weeks (Fig. S14H) post injection. Finally, we confirmed the downregulation of Emc10 in ASO-treated animals via qRT-PCR assays of the HPC, PFC, and SSC brain regions at ten weeks post-treatment (Fig. 6D-F). These results suggest that normalizing Emc10 expression in the brain can ameliorate social and spatial memory deficits in adult Df(16)A+/– mice in a time period of at least two months.

Sustained rescue of cognitive deficits in Df(16)A+/- mice following ASO administration.

(A) Experimental timeline of Emc10ASO2-treated adult Df(16)A+/- mice to determine long-term rescue effect of Emc10 de-repression. This panel was created with BioRender.com. (B-C) Sustained ASO effect on behavioral rescue of Emc10ASO-treated Df(16)A+/- mice in HPC- and PFC-dependent tasks. (B) Rescue of SM deficit in Df(16)A+/- mice after 3- and 8-weeks of ASO injection. After 3-weeks of ASO treatment (left panel), a negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+/- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 28) = 26.20, P<0.0001; post hoc Tukey, P<0.0001]. Increase in the difference score of Df(16)A+/- mice in the Emc10ASO-vs CtrlASO-treated group demonstrates SM rescue (post hoc Tukey, P<0.0001). After 8-weeks of ASO treatment (right panel), a negative difference score (trial 1-trial 2) confirms the SM deficit in CtrlASO-treated adult Df(16)A+/- mice compared to WT littermates [two-way ANOVA for Genotype X Treatment interaction F (1, 28) = 13.36, P=0.0010; post hoc Tukey, P=0.0012. Increase in the difference score of Df(16)A+/- mice in the Emc10ASO- vs CtrlASO-treated group demonstrates prolonged SM rescue (post hoc Tukey, P=0.0007). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+/ mice: n=7, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+/ mice: n=8. (C) Y-maze task displayed rescue of memory impairments in adult Df(16)A+/- male mice 4- and 9-weeks after Emc10ASO injection. Deficits in short-term spatial memory shown by the reduced number of delayed alternations (%) in CtrlASO-treated adult Df(16)A+/- mice can be rescued in Emc10ASO-treated Df(16)A+/- mice after 4-weeks of ASO injection (left panel) [one-way ANOVA, F(3, 29) = 7.585, P=0.0007, post hoc Tukey, P=0.0193]. Deficits in short-term spatial memory in CtrlASO-treated adult Df(16)A+/- mice can be rescued in Emc10ASO-treated Df(16)A+/- mice also after 9-weeks of ASO injection (right panel) [one-way ANOVA, F(3, 29) = 6.547, P=0.0016, post hoc Tukey, P=0.0029]. CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+/ mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+/ mice: n=8. (D-F) Sustained Emc10 de-repression after two months of Emc10ASO injection. ASO-mediated inhibition of Emc10 in different brain regions of Df(16)A+/- mice after 10-weeks of injection. (D) qRT-PCR analysis shows significant upregulation of Emc10 mRNA expression levels in the HPC of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (3, 29) = 13.97, P<0.0001; post hoc Tukey, P=0.0331]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P= <0.0002). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+/ mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+/ mice: n=8. (E) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in the PFC of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (3, 28) = 32.47, P<0.0001; post hoc Tukey, P<0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in the Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P<0.0001). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+/ mice: n=7, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+/ mice: n=8. (F) qRT-PCR analysis shows a significant up-regulation of Emc10 mRNA expression levels in Somatosensory Cortex (SSC) of CtrlASO treated-Df(16)A+/- compared to WT mice [one-way ANOVA, F (3, 28) =23.18, P<0.0001; post hoc Tukey, P=0.0001]. Following ASO treatment, Emc10 expression is normalized to WT levels in Emc10ASO treated-Df(16)A+/– compared to CtrlASO treated-Df(16)A+/- mice (post hoc Tukey, P<0.0001). CtrlASO-treated WT mice: n=9, CtrlASO-treated Df(16)A+/ mice: n=8, Emc10ASO-treated WT mice: n=8 and Emc10ASO-treated Df(16)A+/ male mice: n=7. One- or two-way ANOVA as indicated. Data are presented as mean ± SEM, *P < 0.05, **P < 0.01, ***P < 0.01, ****P<0.0001.

Discussion

Despite an understanding of the molecular mechanisms of 22q11.2DS, especially ones related to abnormal expression and processing of miRNAs (5, 7, 19, 39, 59), we still do not have a promising therapeutic avenue for the cognitive and neuropsychiatric symptoms associated with the 22q11.2 deletion. By leveraging our recent understanding of the molecular, cellular and behavioral consequences of 22q11.2-linked miRNA dysregulation, the present study represents an advancement towards developing a potential therapeutic strategy in two ways:

First, we show that 22q11.2 deletion results in abnormal processing of miRNAs in human neurons and in turn drives upregulation of EMC10 levels as previously described in mouse models (5). Effective reduction to near WT levels or even complete depletion of EMC10 leads to restoration of key alterations in morphological and functional neuronal maturation emerging due to 22q11.2 deletions. The miRNA regulatory mechanism underlying EMC10 upregulation and the restoration of cellular deficits are very similar to the ones we previously described in Df(16)A+/- mice, highlighting the robustness of this molecular alteration as well as the translational value of using animal models to probe the link between Emc10 upregulation and 22q11.2-linked behavioral dysfunction.

Second, we show that normalization of Emc10 levels in adult Df(16)A+/- mouse brain, by ASO-mediated targeted knockdown, is effective in rescuing SM deficits (which emerge during postnatal development and are present as early as postnatal day 22) as well as short-term spatial memory deficits. Significantly, these improvements in cognition were sustained for over two months post-ASO administration. These findings strongly suggest that at least for a subset of 22q11.2-associated cognitive deficits it is the sustained miRNA dysregulation and elevation of Emc10 throughout the adult life that interferes with the underlying neural processes rather than an irreversible impact on brain maturation during early development and demonstrate the therapeutic potential for treating a wide range of cognitive symptoms associated with 22q11.2DS.

In vivo delivery of ASOs offers a potential venue for emerging treatments of genetically driven and postnatally reversible symptoms of neurodevelopmental disorders (NDDs) focusing in reduction of culprit gene expression via sequence-specific knockdown of mRNA transcripts (60). A common challenge in efforts to employ gene-knockdown therapies for dosage-sensitive genes such as EMC10 is restricting target gene expression within optimal levels to avoid potential toxicity due to target hyper-knockdown or complete elimination (61–64). A large number of relatively rare LoF variants or potentially damaging missense variants have been identified in the human EMC10 gene among likely healthy individuals in gnomAD (63, 65) which is depleted of individuals known to be affected by severe NDDs. Taken together with our previous analysis of constitutive heterozygous Emc10 knockout mice (19) that showed no evidence of motor deficits or anxiety-like behavior, these observations strongly suggest that partial reductions in Emc10 levels of approximately 50% are well tolerated at the organismal level and do not result in behavioral abnormalities. The highest dose for both Emc10-specific ASOs used in mutant Df(16)A+/–mice was limited to ∼300 µg and reduced Emc10 mRNA to either normal or below normal (30-50% of WT) expression while attaining full behavioral rescue. While higher dose may be required to ameliorate other behavioral deficits (61, 63), it should be noted that an acute injection of 700µg Emc10ASO in WT mice that resulted in ∼50% reduction of Emc10 levels (Fig. S13) did not cause secondary cellular and behavioral toxicity. Thus, available evidence indicates that therapeutically effective ASO-mediated normalization of EMC10 levels can be up- or down-titrated within a well-tolerated range. However, further work is needed to establish a comprehensive safety profile, including the evaluation of non-cognitive phenotypes, to fully validate the therapeutic potential of Emc10-targeting approaches. Additionally, the long-term effects of Emc10 reduction beyond two months post-injection require further investigation to determine the full extent of its therapeutic benefits. Future studies will also explore whether ASO-mediated normalization of adult Emc10 levels can restore additional 22q11.2 behavioral and cognitive alterations and whether earlier postnatal ASO treatment could prevent the onset of behavioral deficits or mitigate those resistant to adult interventions.

In addition to reduction of Emc10 expression our findings have implications for therapeutic interventions aiming to manipulate its downstream targets. In that respect, it is attractive to speculate that different EMC10 upregulation-linked phenotypes and their developmental requirements may be driven by dysregulation of distinct, individually or in combinations, downstream EMC targets. Such targets are typically multi-transmembrane domain (TMD) proteins (17, 18, 66) that contain low-hydrophobicity TMDs which are hard to insert into ER membrane and thus require the aid of EMC as a membrane insertase (11, 12, 17). Identification of neurotransmitter receptors, channels, and transporters whose biogenesis, trafficking and membrane insertion are affected by EMC10 upregulation could help establish a link between such targets and 22q11.2-related behavioral dysfunction and guide efforts to develop treatments for specific 22q11.2 deletion symptoms.

Overall, by highlighting the manipulation of EMC10 expression and activity, as well as its downstream targets, as a promising alternative or augmentation to currently available treatments, and emphasizing a broad temporal window for therapeutic and preventive intervention in 22q11.2 deletion-associated cognitive and behavioral symptoms, our results lay the foundation for developing mechanism-based therapeutic strategies. These strategies aim to leverage insights from both human and animal models to improve clinical outcomes in precision medicine for neuropsychiatric disorders. Furthermore, our findings may have broader implications for understanding the role of miRNA dysregulation and ER-mediated membrane protein translocation in other neurodevelopmental and neuropsychiatric disorders with overlapping genetic or phenotypic features.

Materials and Methods

Mice

We used Emc10 conditional knockout (see below) and Df(16)A+/- mice (5) in C57BL/6J background. Df(16)A+/- male mice were crossed to C57BL/6J female mice to obtain either Df(16)A+/- or WT littermates. For further details, see Supplemental Materials and Methods.

Generation of Emc10 conditional knockout compound mouse lines

To manipulate the expression of the Emc10 gene in Df(16)A+/–mice we used a Emc10 conditional ‘knockout-first’ mouse design by conducting Cre- and Flip-dependent genetic switch strategies as described earlier (67). This approach enables postnatal manipulation of Emc10 expression at its endogenous locus keeping Emc10 expression within its physiological levels. Emc10tm1a+/ mice (2310044h10rik-Tm1a, MRC Harwell Institute, Oxfordshire, UK) were crossed to a germline Flp mouse line (B6.129S4-Gt(ROSA)26Sortm1(FLP1)Dym/RainJ_JAX:009086) that activates global Flp function and leads to the deletion of the frt-flanked sequence(s) in the offspring. The Emc10tm1c offspring from these cross carries a loxP flanked WT Emc10 allele. We used UBC-Cre/ERT2 mice (B6.Cg-Ndor1 Tg(UBC-cre/ERT2)1Ejb /2J_JAX:008085) in crosses to generate compound Emc10tm1c+/-; UBC-cre/ERT2; Df(16)A+/–mice.

Cell line differentiation

The cell line donors used in this study are listed in Supplementary Table S1. Human iPSC lines were maintained in mTeSR Plus medium on Matrigel coated tissue culture plate and neuronal induction of hiPSC lines into cortical neurons were conducted by using either a combination of small molecule inhibitors or NGN2 overexpression approaches. For further details, see Supplemental Material and Methods.

Genome editing of Q6(22q11.2) hiPCS line

We generated derivatives of the Q6 (22q11.2) hiPSC line carrying either heterozygous (Q6/EMC10HET) or homozygous (Q6/EMC10HOM) EMC10 LoF mutations using standard CRISPR/Cas9 genome editing approaches. For further details, see Supplemental Materials and Methods.

ASOs

Mouse Emc10-targeting ASOs used in these studies were 20 bases in length, chimeric 2’-O-(2-methoxyethyl) (MOE)/DNA) oligonucleotides with phosphodiester and phosphorothioate linkages. The central gap of 10 deoxynucleotides is flanked on its 5’ and 3’ sides by five MOE modified nucleotides. Oligonucleotides were synthesized at Ionis Pharmaceuticals (Carlsbad, CA, USA) as described previously (68, 69). ASOs were solubilized in 0.9% sterile saline or PBS.

Stereotactic Intracerebroventricular (ICV) Injections of ASOs

ASOs were delivered to 8 weeks old mice via ICV injections using a Hamilton syringe (Hamilton Company, Reno, NV, USA) connected to a motorized Stoelting Quintessential Stereotaxic Injector QSI/53311 (Stoelting Co., Wood Dale, IL, USA). For further details, see Supplemental Materials and Methods.

Analysis of dendritic complexity

iNs were prepared on coverslips, fixed at DIV21 and immunostained for TBR1 and MAP2 to identify dendritic branches. For further details, see Supplemental Materials and Methods.

Calcium imaging

Calcium imaging was performed as previously described (70) with a few modifications. For further details, see Supplemental Materials and Methods.

Bulk RNAseq and bioinformatic analysis of mouse hippocampal samples

Total RNA was isolated from 4 WT CtrlASO1, 4 WT Emc10ASO1, 5 Df16 CtrlASO1 and 4 Df16 Emc10ASO1 treated male hippocampi. Stranded polyA+ enriched RNA sequencing libraries were prepared at the Columbia Genome Center (Columbia University, New York, USA) to generate 40 million paired-end reads on Illumina Novaseq 6000 instrument using STRYPOLYA library prep kit. For further details, see Supplemental Material and Methods.

Bulk RNAseq and small RNA/miRNAseq of hiPSC-derived cortical neurons

Total RNA was extracted from hiPSC-derived cortical neurons using the miRVana miRNA isolation kit (#AM1560, Ambion, Thermo Fisher Scientific, Waltham, MA, USA) before bulk RNAseq (paired-ended sequencing; llumina NovaSeq 6000) or miRNAseq (single-end sequencing; Illumina Hiseq 2500) were performed. For further details, see Supplemental Material and Methods.

Behavioral assays

Mice were 11-17 weeks old at the time of behavioral testing except of the social memory assay in juveniles which was performed in 3 week old animals. Behavior was assayed 3-9 weeks following surgical delivery of ASOs or 1 week following TAM/corn oil treatment. The following behavioral assays were performed in this study: Open field assay, social memory assay, contextual fear conditioning and Y-maze delayed alternation task. For further details, see Supplemental Material and Methods. The experimenter was blind to mouse genotype and treatments while performing behavioral assays and data analysis. Animals were given at least one-week intervals between behavioral tests.

Statistical Analysis

Data were analyzed using GraphPad Prism (Graphpad Software, Inc., San Diego, CA, USA). Data were evaluated as indicated, using either unpaired two-tailed t-test, Kolmogorov–Smirnov test, one-way, two-way or three-way analysis of variance (ANOVA) tests followed by post hoc Tukey’s multiple comparison test for comparisons across all groups. Data are presented as mean ±SEM. P values for each comparison are described in the figure legends.